### 2016-03-20 doc/notes/equilerr.tex
add notes

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{hyperref}
\begin{document}

\title{Error of the static flat-distribution sampling}
\author{}
\date{\vspace{-10ex}}
\maketitle

[The content of this note is absorbed
in Secs. II.A.1 and II.F.2]

\section{Introduction}

There are two types of flat-distribution sampling (FDS) methods.
%
The first type of methods,
referred to as the equilibrium FDS (E-FDS) method hereinafter,
fixes the bias potential, $v_i$, during simulation
and only corrects it in the end.
%
Examples include the original
entropic or multicanonical sampling and simulated tempering.
%
The main difficulty is that if a sufficiently accurate bias potential
is unknown before simulation,
it can be hard to achieve the desired flat distribution.

For this reason, the second type of methods,
exemplified by the Wang-Landau (WL) algorithm and metadynamics,
are invented.
%
These methods construct the bias potential on the fly
and achieve a flat distribution in the meantime.
%
Although powerful, they are less well understood,
in that the adaptive updates to the bias potential
breaks the microscopic reversibility
of the underlying sampling mechanism.
%
Thus, for a conservative practitioner,
the E-FDS methods again become appealing
once a sufficiently accurate bias potential is available.
%
Further, with the bias potential fixed to the exact value,
the equilibrium methods can be thought as
the asymptotic limit of the adaptive methods upon convergence,
and thus deliver the maximal possible efficiency.

A practical question now arises.
%
In later stages of an adaptive FDS simulation,
will we be better off switching to the equilibrium FDS
in order to avoid possible efficiency loss due to the adaptive updates?
%
Below, we shall show that it is unnecessary to do so
if the simulation is using the single-bin (WL) updating scheme
with the optimal schedule for controlling the updating magnitude.
%
That is, the bias potential obtained from
an optimal single-bin scheme FDS simulation
will be similar in precision to that from
an E-FDS simulation in long times.
%
Thus, our demonstration below serves as an alternative proof
of the optimality of the single-bin scheme.

\section{Error}

Here is a quick review of E-FDS.
%
In E-FDS, we fix the bias potential, $v_i$, during simulation,
and collect a histogram, $\bar h_i$, via trajectory averaging:
%
\begin{align}
\bar h_i
=
\frac 1 t
\sum_{ t_1 = 1 }^{ t }
  h_i( t_1 )
=
\frac 1 t
\sum_{ t_1 = 1 }^{ t }
  \delta_{ i, i(t_1) },
\label{eq:hbar_def}
\end{align}
%
where $h_i(t_1) = \delta_{ i, i(t_1) }$
is the instantaneous histogram.
%
In the end,
we use this histogram to correct the bias potential as
%
\begin{align}
v_i^\mathrm{corr}
=
v_i
+
\log \frac { \bar h_i }
           {      p_i }.
\end{align}

The error of the method depends on the precision of the histogram.
%
Below we shall evaluate the error in the ideal case
in which $v_i$'s assume the exact values.
Then,
%
\begin{align}
E
=
\left\langle
  \sum_{ i = 1 }^ n
    p_i \,
    \left(
      \log \frac { \bar h_i }
                 {      p_i }
    \right)^2
\right\rangle
\approx
\sum_{ i = 1 }^ n
  p_i \,
  \left\langle
    \left(
      \frac { \bar h_i }
            {      p_i }
      - 1
    \right)^2
  \right\rangle.
\end{align}

To make better comparison with
the error of the single-bin case,
we shall rewrite the error as
a sum of the errors of the eigenmodes.
We define
\begin{align}
\bar \eta_k
=
\frac{ 1 } { t }
\sum_{ t_1 = 1 } ^ t
  \eta_k( t_1 ),
\end{align}
%
with
\begin{align}
\eta_k( t_1 )
=
\sum_{ i = 1 }^ n
\phi_{k i} \left[
             \frac{ h_i(t_1) }
                  { p_i      }
             - 1
           \right],
\end{align}
and
\begin{align}
\sum_{ k = 1 }^n
  \phi_{k i} \, \phi_{k j}
  = p_i \, \delta_{i j}.
\end{align}
%
Then,
%
\begin{align}
E
=
\sum_{ k = 1 }^n
  \left\langle
    \bar \eta_k^2
  \right\rangle,
\label{eq:error_sbin_equil}
\end{align}
%
because
%
\begin{align*}
\sum_{ k = 1 }^n
  \bar \eta_k^2
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ k = 1 }^n
    \eta_k( t_1 ) \, \eta_k( t_2 )
\\
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ i, j = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
    \left[
      \frac{ h_j(t_2) }
           { p_j      }
      - 1
    \right]
    \sum_{ k = 1 }^n
      \phi_{k \, i} \, \phi_{k \, j}
\\
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ i, j = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
    \left[
      \frac{ h_j(t_2) }
           { p_j      }
      - 1
    \right]
    p_i \, \delta_{i, j}
\\
&=
\sum_{ i = 1 }^n
\frac{ 1 } { t }
\sum_{ t_1 = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
\frac{ 1 } { t }
\sum_{ t_2 = 1 }^n
    \left[
      \frac{ h_i(t_2) }
           { p_i      }
      - 1
    \right]
\\
&=
\sum_{ i = 1 }^n
  p_i \,
    \left(
      \frac{ \bar h_i }
           { p_i      }
      - 1
    \right)^2
.
\end{align*}
%
The sum can start from $k = 2$,
because by $\phi_{1i} = p_i$,
we have
$$
\eta_1(t_1)
=
\sum_{ i = 1 }^n
  \phi_{1i} \, \frac{ h_i(t_1) - p_i } { p_i }
=
\sum_{ i = 1 }^n h_i(t_1)
- \sum_{ i = 1 }^n p_i = 0.
$$

Finally, we shall evaluate Eq. \eqref{eq:error_sbin_equil}
in the long-time limit:
\begin{align*}
\left\langle
  \bar \eta_k^2
\right\rangle
&=
\frac{1}{t^2}
\sum_{ t_1, t_2 = 1 }^n
\langle \eta_k(t_1) \, \eta_k(t_2) \rangle
\\
&\approx
\frac{1}{t}
\sum_{ \tau = -\infty }^{ \infty }
\langle \eta_k(t_1) \, \eta_k(t_1 + \tau) \rangle
\\
&=
\frac{1 }{t}
\int_{-\infty}^\infty
\kappa_n(\tau) \, d\tau
=
\frac{1}{t}
\Gamma_k.
\end{align*}
%
Thus,
\begin{align}
E
=
\frac{ 1 } { t }
\sum_{ k = 2 }^n
\Gamma_k
\end{align}

Now comparing this with the error of the single-bin scheme
with the optimal schedule,
$$
E_{\mathrm{sb}}
=
\frac{    t    }
     { (t + t_0)^2 }
\sum_{ k = 2 }^n
  \Gamma_k,
$$
we find that the difference is negligible in the long time limit.

\end{document}