### 2017-02-01 doc/old/V4.1_jd/invtV4.1_jd.tex
invt.tex: add the Justin edit

\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

\hypersetup{
  colorlinks,
  linkcolor={red!30!black},
  citecolor={green!20!black},
  urlcolor={blue!80!black}
}

\definecolor{DarkBlue}{RGB}{0,0,64}
\definecolor{DarkBrown}{RGB}{64,20,10}
\definecolor{DarkGreen}{RGB}{0,64,0}
\definecolor{DarkPurple}{RGB}{64,0,42}
\definecolor{LightGray}{gray}{0.85}
\newcommand{\repl}[2]{{\color{gray} [#1] }{\color{blue} #2}}
\newcommand{\add}[1]{{\color{blue} #1}}
\newcommand{\del}[1]{{\color{gray} [#1]}}
\newcommand{\note}[1]{{\color{DarkGreen}\footnotesize \textsc{Note.} #1}}
\newcommand{\answer}[1]{{\color{DarkBlue}\footnotesize \textsc{Answer.} #1}}
\newcommand{\summary}[1]{{\color{DarkPurple}\footnotesize \textsc{Summary.} #1}}

\newcommand{\Err}{E}
\newcommand{\ii}{\mathrm{i}}

\begin{document}

\title{Optimal updating factor in adaptive flat-distribution-sampling simulations}

\author{Cheng Zhang}
\author{Justin A. Drake}
\affiliation{
Sealy Center for Structural Biology and Molecular Biophysics,
The University of Texas Medical Branch,
Galveston, Texas 77555-0304, USA}
\author{Jianpeng Ma}
\affiliation{
Department of Biochemistry and Molecular Biology,
Baylor College of Medicine, Houston, Texas 77030, USA}
\affiliation{
Department of Bioengineering,
Rice University, Houston, Texas 77005, USA}
\author{B. Montgomery Pettitt}
\email{bmpettitt@utmb.edu}
\affiliation{
Sealy Center for Structural Biology and Molecular Biophysics,
The University of Texas Medical Branch,
Galveston, Texas 77555-0304, USA}

\begin{abstract}
  We present a method of computing the optimal schedule
  of the updating magnitude
  for a class of free energy calculation methods including
  the Wang-Landau (WL) algorithm and metadynamics.
  These methods rely on adaptive construction of
  a bias potential that offsets
  the potential of mean force by histogram-based adaptive updates.
  The updating magnitude should decrease over time
  to reduce the noise in the bias potential,
  and one may speed up the convergence by choosing an optimal schedule
  of the updating magnitude.
  We will show that
  the optimal schedule depends on the updating scheme (WL or metadynamics),
  and the initial updating magnitude should be half of
  the previous equilibrium value.
  Further,
  the WL updating scheme
  belongs to a class of bandpass updating schemes
  that are optimal for asymptotic convergence,
  and the optimal schedule of these schemes
  is the inverse number of sampling steps with a proper shift,
  as given by the $1/t$ prescription.
\end{abstract}

\maketitle

\section{Introduction}

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry that can provide insight into
an array of phenomena not easily studied with traditional experiments.
Given a system,
often the task is to compute
a distribution, $p^*(z)$,
along a collective variable $z$ by sampling with Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
The free energy, or potential of mean force (PMF), along the collective variable is proportional to $\ln p^*(z)$.
However, accurately estimating the PMF is often difficult when the collective variable is energetically restricted to local regions of the complex, multi-modal free energy surface.
Thus,
to overcome these energetic barriers and to capture the global shape of the free energy surface
an effective strategy is to introduce a bias potential that
cancels the PMF. Ideally, then, sampling yields a flat distribution
along $z$\cite{mezei1987, berg1992, lee1993,
wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.

The Wang-Landau (WL) algorithm\cite{wang2001, *wang2001pre}
and metadynamics\cite{huber1994, laio2002,
laio2008, barducci2011, sutto2012}
are two widely-used and closely-related\cite{micheletti2004}
flat-distribution-sampling (FDS) techniques
that build the bias potential associated with the collective variable adaptively, or on-the-fly, during simulation.
Both techniques regularly update the bias potential
to discourage future visits to previously sampled configurations
by incrementally elevating the bias potential.
The difference lies
in the updating window function,
which takes the form of a discrete
$\delta$-function (or a boxcar function one histogram bin wide)
in WL ,
or that of a Gaussian
in metadynamics.
In both cases,
regular updates to the bias potential
disrupts the underlying equilibrium
sampling\cite{laio2002},
and for convergence of the PMF
one  has to gradually decrease
the updating magnitude of the bias.

The schedule of reducing
the updating magnitude,
therefore, critically affects
the precision of the final bias potential, and
hence that of the PMF\cite{liang2007,
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, *caparica2014,
barducci2008, dickson2011, dama2014}.
For the WL algorithm, the optimal updating schedule
is a function of the
inverse time\cite{liang2007,
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
morozov2007, zhou2008}
or
the inverse number of simulation steps.
The optimal schedule for metadynamics,
however, is less studied.

In this study,
we present a general formalism for computing
the optimal schedule
for a any adaptive FDS method,
including the WL algorithm and metadynamics.
Our method maps the optimization problem to a mechanical one,
in which the schedule plays the role of the velocity of
a free particle with a position-dependent mass,
and the final error becomes the action.
In this way, the optimal schedule
can be deduced from the equation of motion
of the particle that minimizes the action.
We show that the optimal schedule
depends on how the bias potential is updated,
and hence is different in the WL and metadynamics cases.
While the optimal schedule in the WL case
is given by the known inverse-time formula,
the optimal schedule for metadynamics is more complex
and sensitive to the simulation length
and the width of the Gaussian window.
In either case, however,
the initial updating magnitude
should be set to
half of the previous equilibrium value,
which amounts to a shift of the origin of time in the WL case.
Further, we can generalize
the WL updating scheme
to a class of bandpass updating schemes
that are optimal for long simulations,
and the above origin-shifted inverse-time schedule
turns out to be optimal for these schemes as well.
These results may help to better
plan and optimize FDS simulations, which, at present, is done heuristically and through trial-and-error.

The article is organized as follows.
We present the analytical results in Sec. \ref{sec:theory},
numerically verify some aspects
in Sec. \ref{sec:results},
and conclude the article
in Sec. \ref{sec:conclusion}.

\section{\label{sec:theory}
Theory}

We develop the theory
in the following order.
In Sec. \ref{sec:background},
we first review the basics of FDS
and some known aspects of the optimal schedule.
Then, in Sec. \ref{sec:single-bin},
we illustrate the method of
computing the optimal schedule
in the simple case of the
single-bin updating scheme
used by the WL algorithm,
by proving the optimality
of the known inverse-time formula.
Next, we extend the method to the general case
of a multiple-bin updating scheme
that encompasses metadynamics
in Secs. \ref{sec:multiple-bin}
to \ref{sec:band-matrix}.
Finally, we compare different updating schemes
and show that the single-bin scheme
and a class of generalized bandpass schemes
are optimal for convergence
in the long time or extensive sampling limit
in Sec. \ref{sec:cmpschemes}.

\subsection{\label{sec:background}
Background}

\subsubsection{\label{sec:FDS}
Flat-distribution sampling}

Consider the problem of computing
the distribution, $p^*_i$,
along a discrete quantity, $i$,
for a given system.
For example, $i$ can be the energy, $E$,
in a lattice spin model or the temperature index
in a simulated tempering simulation\cite{
marinari1992, lyubartsev1992}.
For a continuous quantity, $z$,
which is often the case in a (classical) molecular system,
we can discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin,
$(z, z + \Delta z)$.
The distribution is normalized as
$\sum_{i = 1}^n p^*_i = 1$.

For a large system at temperatures of interest,
the distribution $p^*_i$ often tends to
be localized,
and to explore the global shape
of the PMF, $-\ln p^*_i$,
it is often advantageous to perform
biased sampling that generates a
a wider, smoother target distribution, $p_i$.

To do so, we introduce a bias potential, $V_i$,
which changes the equilibrium distribution to
\begin{equation}
  \pi_i \propto p^*_i \, e^{-V_i}.
  \label{eq:pi_p_V1}
\end{equation}
Upon normalization,
$\sum_{i = 1}^n \pi_i = 1$,
we get
\begin{equation}
  \pi_i
  =
  \frac{                p^*_i \, e^{-V_i} }
       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
  .
  \label{eq:pi_p_V}
\end{equation}

We wish the biased equilibrium distribution, $\pi_i$,
to converge to the target distribution, $p_i$,
which is usually chosen to be flat
(i.e. $p_i$ constant for all $i$'s).\cite{
dayal2004, trebst2004, zhang2007, barducci2008, singh2011}.
We will generally call such a biased simulation
a flat-distribution-sampling (FDS) simulation.
To fulfill this goal, the bias potential must satisfy
\begin{equation}
  V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
  ,
  \label{eq:Vi_target}
\end{equation}
according to Eq. \eqref{eq:pi_p_V1}.
This relation thus sets a practical goal for the convergence of FDS
methods---adjusting the bias potential to make the
observed distribution, or the normalized histogram,
match the intended one.
Once this aim is achieved,
the PMF, $-\ln p^*_i$, can be naturally
deduced from the bias potential. % $V_i$.

Roughly speaking, there are two types of FDS methods.
The first type,
referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
uses a fixed bias potential, $V_i$ (estimated a priori),
during simulation.
Examples include the original
entropic or multicanonical sampling\cite{
berg1992, lee1993}
and simulated tempering\cite{
marinari1992, lyubartsev1992}
with $i$ being the index of energy
and that of temperature, respectively.
The estimated bias potentials at bin $i$, $V_i$,
can be improved
with the normalized histogram of the sampling, $H_i$, accumulated
during simulation.
Using $H_i$ as an estimate of $\pi_i$ in
Eq. \eqref{eq:pi_p_V1}
allows us to correct $V_i$ as
\begin{equation}
  V^\mathrm{corr}_i
  =
  V_i
  +
  \ln \frac{ H_i }
           { p_i }.
  \label{eq:vcorr_equil}
\end{equation}
However, for an unknown system,
it is often difficult
to guess a sufficiently accurate
$V_i$ with which to start.
Thus,
it can take many iterative rounds of simulations
to achieve the desired flat distribution.
The effectiveness of this approach
depends on the precision of the histogram, $H_i$,
which in turn depends on the simulation length
that must be decided in advance.
Thus, the rigidity of this framework disallows
continuous, rapid improvement of the bias potential, and subsequently the PMF,
as simulation lengthens.

The second type of methods,
referred herein as adaptive FDS methods,
address the above shortcomings
by continuously and incrementally updating the bias potential to accelerate convergence of the target distribution.
Examples include
the WL algorithm and metadynamics.
However,
as mentioned, the on-the-fly updates break the microscopic reversibility
of the underlying sampling mechanism.
Further, failure to properly reduce the magnitude of
the histogram-based increments
would lead to roughness or errors in the final estimate of the PMF.
To minimize this effect,
many adaptive FDS methods adopt a schedule to
reduce the updating magnitude
over the course of simulation\cite{
marsili2006,
liang2007,
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
barducci2008}.
For example, one can show that
if the magnitude decreases according to
the optimal inverse-time prescription\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, *caparica2014},
the WL algorithm is equivalent to the ideal equilibrium counterpart
in improving the precision of the bias potential, thus
allowing for maximal efficiency
(cf. Appendix \ref{sec:equilerr}).

\subsubsection{Updating schemes}

We will mainly focus on the adaptive methods below.
In the WL algorithm\cite{wang2001, *wang2001pre},
the bias potential, $V_i(t)$, is updated
at each sampling step $t$ as
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i }
  ,
\label{eq:wl_update}
\end{equation}
where $i(t)$ is the bin at step $t$,
and $\alpha(t)$ is the updating magnitude.
We refer to this scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$.
In a more general \emph{multiple-bin scheme}, however,
several neighboring bins are updated:
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) }
       { p_{i(t)} },
  \label{eq:mbin_update}
\end{equation}
where the elements, $w_{ij}$'s, are a set of coefficients satisfying
the restrictions to be discussed
in Sec. \ref{sec:updating-matrix}.
The Gaussian updating scheme
used in metadynamics
is a multiple-bin scheme.

One can show that with a constant magnitude $\alpha(t) = \alpha > 0$,
the distribution, $\pi_i$, collected from
an adaptive FDS simulation
is identical to $p_i$ in the long time limit.
However, as the above updates
change $V_i(t)$ continuously,
the equilibrium distribution ()Eq. \eqref{eq:pi_p_V}) no longer holds
at all times,
and the error in $V_i(t)$
depends on the updating magnitude\cite{
  zhou2005, liang2005, laio2005, bussi2006, poulain2006, liang2007,
  morozov2007, zhou2008, morozov2009, crespo2010,
  atchade2011, fort2015}.
Thus, one has to reduce the magnitude $\alpha$ over time
to improve the estimate of the distribution,
$p^*_i$. %via Eq. \eqref{eq:v_def}.

\subsubsection{Updating magnitude and the inverse-time formula}

In the original WL papers\cite{
wang2001, *wang2001pre},
the updating magnitude $\alpha$
(or $\ln f$ referred therein)
is controlled in stages.
In each stage, $\alpha$
is kept constant,
and the histogram along $i$
is collected and monitored.
Once the histogram is sufficiently flat,
we switch to a new stage
with $\alpha$ reduced by a factor of $1/2$\cite{
wang2001, *wang2001pre}.
While this scheme works well for early stages,
it tends to reduce $\alpha$
too quickly in late stages, making the asymptotic error
saturate\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008}.

Later,
it was found that
a better way
of controlling the schedule $\alpha(t)$
in late stages
is to follow the formula
\begin{equation}
  \alpha(t) = \frac{1}{t},
  \label{eq:alpha_invt}
\end{equation}
where $t$,
referred to as the ``time'' below,
is the number of steps
from the beginning of the simulation.
This simple yet effective formula has attracted
several studies\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, *caparica2014}.
Here we will re-derive this formula
and suggest a slight modification
whereby the origin of $t$ is shifted
according to the initial error.
Interestingly, we may reach the same result
from optimally planning the simulation lengths in the stages
and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.

As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
several similar ways of
reducing the updating magnitude
were developed for umbrella sampling and metadynamics,
particularly, the self-healing umbrella sampling
and well-tempered metadynamics\cite{
  marsili2006, barducci2008, dickson2011, dama2014}.
Our aim here is to find the optimal schedule $\alpha(t)$,
for a general, universal updating scheme.

\subsection{\label{sec:single-bin}
Single-bin updating scheme}

In this section,
we will derive the optimal $\alpha(t)$
for the single-bin scheme,
Eq. \eqref{eq:wl_update}.
To do so,
we will first express the error of
the bias potential, $V_i(t)$,
as a functional of $\alpha(t)$,
and then minimize it by variation.

\subsubsection{Error}

First, we need to define the error in the bias potential.
This is done by removing two contributions from the bias potential.
The first contribution comes from the target value,
Eq. \eqref{eq:Vi_target}.
For this purpose,
we introduce a shifted bias potential
\begin{equation}
  v_i(t)
  \equiv
  V_i(t)
  -
  \ln \frac { p^*_i }
            { p_i }
  ,
  \label{eq:v_def}
\end{equation}
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i$'s are
equivalent to the updates to $V_i$'s,
since the shift
remains a constant during the course of simulation,

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the
resulting distribution $\pi_i(t)$,
the real deviation is more readily defined
from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
that represents the average growth\cite{
dama2014},
\begin{equation}
  x_i(t) \equiv v_i(t) - \bar v(t)
  ,
\label{eq:x_def}
\end{equation}
where
\begin{equation}
  \bar v(t)
  = \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vbar_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$.
In this way,
the variable, $x_i(t)$,
represents the error of the bias potential,
and is expected to vanish on convergence.
We, therefore, define
the error of the bias potential as
\begin{equation}
  \Err
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_sum}
\end{equation}

Note that the distribution, $\pmb \rho$,
introduced above is independent of
the target distribution $\mathbf p$.
For example,
in well-tempered metadynamics,
$\pmb\rho$ is a flat distribution,
while $\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
barducci2008, dama2014}.
From Eqs. \eqref{eq:x_def} and \eqref{eq:vbar_def},
we have
\begin{equation}
  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
  .
\label{eq:rhox_sum}
\end{equation}

\subsubsection{Differential equation}

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation
%
\begin{equation}
  \dot V_i(t)
  \equiv
  \frac{ dV_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i },
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise.
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.

Next, we split $h_i(t)$ into a deterministic averaged part
and a stochastic part, $\zeta_i(t)$,
\begin{equation}
  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
The deterministic part can be related
to an ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$ and
the same bias potential at time $t$.
The initial states and the stochastic forces
during the process may differ, however.

For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period.
Then,
the sampling process is approximately Markovian, % of a finite order,
and we may assume Eq. \eqref{eq:pi_p_V}
for the deterministic part
\begin{align}
  \langle h_i(t) \rangle
  &\approx
  \pi_i(t)
  =
  \frac{                p_i \, e^{-x_i(t)} }
       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
  \notag
  \\
  &\approx
  p_i \,
  \left[
    1 - x_i(t) + \sum_{r=1}^n p_r \, x_r(t)
  \right]
  ,
  \label{eq:h_ave}
\end{align}
where we have assumed the smallness
of $x_i(t)$ in the linear expansion
and used $\sum_{j=1}^n p_j = 1$
on the second line.

For the random noise part,
the average is zero by definition,
$\langle \zeta_i(t) \rangle = 0$,
and the autocorrelation functions
depend only on the time difference
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \sigma_{ij}(t - t'),
  \notag
\end{equation}
where $\sigma_{ij}(t)$ is an even function
that vanishes at large $t$.

From Eqs.
\eqref{eq:v_def}-\eqref{eq:vbar_def},
\eqref{eq:rhox_sum}-\eqref{eq:h_ave},
we get a set of decoupled equations
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
\label{eq:dxdt_WL}
\end{equation}
where
\begin{equation}
  \xi_i(t)
  =
  \frac{ \zeta_i(t) }
       { p_i }
  -
  \sum_{ r = 1 }^n
    \rho_r
    \frac{ \zeta_r(t) }
         {     p_r    }
  .
\label{eq:xi_def}
\end{equation}

The total error defined in Eq. \eqref{eq:error_sum}
now depends implicitly on the function $\alpha(t)$
via $x_i(t)$.
Our aim is to find the $\alpha(t)$
that minimizes the error
at the end of the simulation period $T$.
It suffices to consider
a one-variable problem that
minimizes each individual
$\left\langle x_i^2(T) \right\rangle$.

\subsubsection{\label{sec:onevar}
Optimization of a one-variable problem}

Consider the following stochastic differential equation
of a single variable:
\begin{equation}
\dot x(t) = -\alpha(t) \, \left[ x(t) - \xi(t) \right],
\label{eq:dxdt_alpha}
\end{equation}
where $\xi(t)$ is a generalized noise
that is invariant under time translation:
$
  \left\langle \xi(t) \, \xi(t') \right\rangle
  =
  \kappa(t - t')
  .
$
Additionally,
we assume that $\kappa(t)$ is an even function and
\begin{equation}
  \lim_{t \rightarrow \infty} \kappa(t)
  = 0
  .
  \label{eq:kappa_limit}
\end{equation}
For example, for a white noise,
$\kappa(t)$ is proportional to a
Dirac $\delta$-function, $\delta(t)$.
We wish to show that the $\alpha(t)$
of minimizing $\langle x^2(t) \rangle$ at long times
is given by
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }.
\label{eq:alpha_invt1}
\end{equation}

To do so, we first recall
the formal solution of Eq. \eqref{eq:dxdt_alpha}:
%
\begin{equation}
x(T)
=
x(0) \, e^{-q(T)}
+
\int_0^T
  \dot u\bigl( q(t) \bigr) \, \xi(t) \, dt,
\notag
\end{equation}
%
where
%
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
%
and
%
\begin{align}
u(q')
&\equiv
e^{q' - q(T)}.
\label{eq:u_def}
\end{align}
At the end of period $T$, we have
\begin{align}
&\left\langle x^2(T) \right\rangle
=
\langle x^2(0) \,
\rangle e^{-2 \, q(T)}
\notag
\\
&\qquad
+
\int_0^T \int_0^T
  \dot u\bigl( q(t) \bigr) \,
  \dot u\bigl( q(t') \bigr) \,
  \kappa(t - t') \, dt' \, dt.
\label{eq:x2t_average}
\end{align}

This is a functional of the curve $q(t)$.
We now fix the endpoint value $q(T)$
and vary the curve, $q(t)$ ($0 < t < T$).
This is equivalent to varying $u$
with fixed endpoints,
$u\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$u\bigl( q(T) \bigr) = 1$.
Note also that the first term on the right-hand side
of Eq. \eqref{eq:x2t_average} is fixed during
the process.
By the Euler-Lagrange equation, we get
$$
\begin{aligned}
0
&=
\frac{d}{dt} \int_0^T
  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
\\
&= \int_0^T
  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by assuming Eq. \eqref{eq:kappa_limit}
and a large $T$.
In order for this to hold for any $t$,
we must have
$
\ddot u\bigl( q(t') \bigr) = 0
,
$
or
\begin{equation}
  \dot u\bigl( q(t') \bigr) = c
  ,
\label{eq:dudt_const}
\end{equation}
where $c$ is a constant of $t'$.
Using Eq. \eqref{eq:u_def},
we get
$$
e^{ q(t') - q(T) }
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
Taking the logarithm,
and differentiating both sides
with respect to $t'$
yields Eq. \eqref{eq:alpha_invt1}.

\subsubsection{Optimal schedule for the single-bin scheme}

For the problem of
minimizing the error defined in Eq. \eqref{eq:error_sum}
for the single-bin scheme,
we note that Eq. \eqref{eq:dxdt_WL}
is of the form of Eq. \eqref{eq:dxdt_alpha}.
Thus,
the optimal schedule $\alpha(t)$ is given by
Eq. \eqref{eq:alpha_invt1}
for it optimizes each term $\langle x_i^2 \rangle$,
and hence the sum;
Eq. \eqref{eq:alpha_invt}
is the special case of $t_0 = 0$.
In Sec. \ref{sec:finlencorr},
we show that % $t_0$ can be adjusted
one can further improve the schedule by adjusting $t_0$
to accommodate the initial error.

\subsection{\label{sec:multiple-bin}
Multiple-bin updating scheme}

We now consider a multiple-bin updating scheme,
given in Eq. \eqref{eq:mbin_update}.
In this case,
upon a visit to bin $j$,
we update not only the bias potential at bin $j$,
but also those at a few neighboring bins, $i$'s.
The relative updating magnitudes,
$w_{ij}$'s,
form a matrix $\mathbf w$.
Then, Eq. \eqref{eq:vt_diffeq} is generalized to
\begin{equation}
  \dot V_i(t)
  =
  \alpha(t) \sum_{j=1}^n \, w_{ij} \frac{ h_j(t) } { p_j }.
  \label{eq:vt_diffeq_mbin}
\end{equation}

The optimization procedure is similar to the single-bin case.
However, we will assume some restrictions and approximations
to simplify the discussion.

\subsubsection{\label{sec:updating-matrix}
Updating matrix}

The matrix $\mathbf w$ is not arbitrary.
First, to sample
the desired distribution
$\mathbf p = (p_1, \dots, p_n)$,
 $\dot V_i(t)$
in Eq. \eqref{eq:vt_diffeq_mbin}
should be a constant of $i$
if $h_j(t)$ were the same as $p_j$.
This allows the bias potential to grow uniformly
in the asymptotic regime.
Thus, by a proper scaling of $\alpha(t)$,
we may write this condition as
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1
  .
\label{eq:w_sumj}
\end{equation}
In other words, $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
except that some elements can be negative.
Below we will take advantage of the analogy
and borrow techniques used
in studying transition matrices\cite{vankampen}.

To simplify the ensuing discussion,
we  limit ourselves to matrices % $\mathbf w$'s
that satisfy
the detailed balance condition,
\begin{equation}
  \rho_i \, w_{ij} = \rho_j \, w_{ji}
  ,
  \label{eq:w_detailedbalance}
\end{equation}
for the distribution $\pmb \rho$
introduced in Eq. \eqref{eq:vbar_def}.
It follows that
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j,
  \label{eq:w_balance}
\end{equation}
i.e., $\mathbf \rho$ is a left eigenvector of
$\mathbf w$ with eigenvalue $1$.

By assuming Eq. \eqref{eq:w_detailedbalance},
we can diagonalize $\mathbf w$ with a set of
eigenvectors, $\phi_{ki}$,
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij}
  =
  \lambda_k \, \phi_{kj}
  ,
\label{eq:eig_w}
\end{equation}
satisfying the orthonormal conditions\cite{vankampen}:
\begin{align}
  \sum_{k = 1}^n
    \phi_{ki} \, \phi_{kj}
  &=
  \delta_{ij} \, \rho_i,
  \label{eq:eig_orthonormal_cols}
  \\
  \sum_{i = 1}^n
    \frac{ \phi_{ki} \, \phi_{li} }
         { \rho_i }
  &=
  \delta_{kl}
  .
\label{eq:eig_orthonormal_rows}
\end{align}

In particular, there is always an eigenvector
corresponding to the distribution, $\pmb \rho$.
We will label this eigenvector as the first one,
and
\begin{equation}
  \phi_{1i} = \rho_i,
\label{eq:eigenmode1}
\end{equation}
which satisfies Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 1$.
The corresponding eigenvalue is $\lambda_1 = 1$
by Eq. \eqref{eq:w_balance}.
Further, for $k \ge 2$, we have
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \sum_{ i = 1 }^n
    \frac{ \phi_{ki} \, \phi_{1i} }
         { \rho_i }
  =
  0
  .
\label{eq:ortho1}
\end{equation}

Finally, there is a stability condition.
In a \emph{stable} updating scheme,
no eigenmode grows indefinitely,
which means that no eigenvalue $\lambda_k$
can be negative.

\subsubsection{Linearization and decoupling to eigenmodes}

We can now simplify Eq. \eqref{eq:vt_diffeq_mbin}.
By using Eqs. \eqref{eq:vbar_def}
and \eqref{eq:w_balance},
we can show that Eq. \eqref{eq:dvbardt}
remains true,
and by using Eqs.
\eqref{eq:x_def},
\eqref{eq:rhox_sum},
\eqref{eq:h_ave},
\eqref{eq:xi_def},
and \eqref{eq:w_sumj}, we get
$$
\begin{aligned}
  \dot x_i(t)
  &= \alpha(t) \sum_{j=1}^n w_{ij}
  \left(
    \frac{ h_j(t) } { p_j }
    -
    \sum_{r = 1}^n
      \rho_r \frac{ h_r(t) } { p_r }
  \right)
  \\
  &\approx
  -\alpha(t) \sum_{j = 1}^n
  w_{ij} \left[ x_j(t) - \xi_j (t) \right]
  .
\end{aligned}
$$

Diagonalizing the matrix $\mathbf w$
using the eigenvectors defined in Eqs.
\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows},
we get a set of decoupled equations
\begin{equation}
  \dot y_k(t)
  =
  -\alpha(t) \, \lambda_k \, [y_k(t) - \theta_k(t)],
\label{eq:yt_diffeq}
\end{equation}
where
\begin{align}
  y_k(t)
  &=
  \sum_{j=1}^n
    \phi_{kj} \, x_j(t)
  ,
  \label{eq:y_def}
\end{align}
and
\begin{align}
  \theta_k(t)
  &=
  \sum_{j=1}^n
    \phi_{kj} \, \xi_j(t)
  ,
  \label{eq:theta_def}
\end{align}
are the deviation and noise of mode $k$, respectively.

\subsubsection{Error function}

The error function defined in Eq. \eqref{eq:error_sum}
can be rewritten in terms of the eigenmodes of $\mathbf w$
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 1}^n
    \left\langle
      y_k^2
    \right\rangle
  ,
\notag
\end{align}
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
From Eq. \eqref{eq:eigenmode1},
the first eigenmode,
$y_1 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
from $k = 2$,
\begin{equation}
  \Err
  =
  \sum_{k = 2}^n
    \left\langle
    y_k^2(T)
    \right\rangle
  .
\label{eq:y2_sum}
\end{equation}

In analogy to Eq. \eqref{eq:x2t_average},
using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum}, we can write
the error after a period $T$
as a sum of two contributions,
\begin{equation}
  \Err
  =
  \Err_R + \Err_A,
  \label{eq:error_split}
\end{equation}
where the residual error
(from the decay of the initial error)
is given by
%
\begin{equation}
  \Err_R
  =
  \sum_{k = 2}^n
    \left\langle y_k^2(0) \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
and the asymptotic error
is given by
%
\begin{equation}
  \Err_A
  =
  \int_0^T \int_0^T
  \sum_{k = 2}^n
    \dot u_k\bigl( q(t) \bigr) \,
    \dot u_k\bigl( q(t') \bigr) \,
    \kappa_k(t - t') \, dt \, dt',
  \label{eq:error_asym}
\end{equation}
%
with
%
\begin{align*}
  u_k(q')
  &\equiv
  e^{\lambda_k \, [q' - q(T)]},
  \\
  \kappa_k(t - t')
  &\equiv
  \left\langle
    \theta_k(t) \, \theta_k(t')
  \right\rangle.
\end{align*}

\subsubsection{White-noise approximation}

In the asymptotic regime
(i.e., if the simulation period, $T$, is sufficiently long),
we can approximate $\kappa_k(t)$ as
%
\begin{equation}
  \kappa_k(t) \approx \Gamma_k \, \delta(t)
  ,
\label{eq:kappa_delta}
\end{equation}
%
where
%
\begin{equation}
  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
  \notag
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
  \label{eq:Gamma_sum}
\end{equation}
Then, Eq. \eqref{eq:error_asym} is simplified as
\begin{align}
  \Err_A
  &=
  \int_0^T
  \sum_{k = 2}^n
  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
\label{eq:error_asym1}
\end{align}
In this way,
the underlying sampling process
affects the error only through the
few $\Gamma_k$'s.
In Appendix \ref{sec:Gamma},
we evaluate $\Gamma_k$'s in two special cases,
and show how to measure them
from an adaptive FDS simulation.

\subsection{\label{sec:mbin_opta}
Optimal schedule for multiple-bin updating schemes}

\subsubsection{\label{sec:eqlerr}
Equilibrium error under a constant updating magnitude
}

Before addressing the optimal schedule
for a multiple-bin scheme,
we will first compute the equilibrium error
under a constant schedule,
$\alpha(t) = a_0$.
After a long period $T'$,
the residual error becomes negligible, and
the $k$th component of the asymptotic error
is given by
\begin{align}
  \left\langle
    y_k^2
  \right\rangle
  &=
  \int_0^{T'}
    \Gamma_k \, (\lambda_k \, a_0)^2 \,
      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
    \, dt
  \notag
  \\
  &=
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  ,
  \qquad
  ( T' \to \infty )
  .
\label{eq:y2_eql}
\end{align}

We will use this result to model the initial error
before starting the schedule $\alpha(t)$ to be optimized.
That is, we assume
$\alpha(t) = a_0$
for $t < 0$, and
the system has entered the sampling equilibrium
under a fixed $a_0$ at $t = 0$.
We can then use Eq. \eqref{eq:y2_eql}
for the $\langle y_k^2(0) \rangle$'s
in Eq. \eqref{eq:error_res},
and
the residual error at the end of period $T$
is given by
\begin{equation}
  \Err_R
  =
  \sum_{k = 2}^n
      \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k \,
      e^{  - 2 \, \lambda_k \, q(T) }
  .
  \label{eq:error_res1}
\end{equation}
Note that the residual error
depends on $\alpha(t)$ only through
the integral $q(T) = \int_0^T \alpha(t) \, dt$.
Thus, if we vary $\alpha(t)$ under a fixed $q(T)$,
this residual part can be ignored,
and the optimization can be done exclusively
on the asymptotic part, $\Err_A$.

\subsubsection{\label{sec:optschedule}
Optimal schedule}

As shown above, with a fixed value of $q(T)$,
the optimization of the total error is equivalent to
that of the asymptotic part.
We can rewrite Eq. \eqref{eq:error_asym1} much like an action
of a particle whose position is given by $q(t)$:
\begin{equation}
  \Err_A
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
\label{eq:error_asym_Lagrangian}
\end{equation}
where the Lagrangian is
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 2 }^n
    \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
  =
  \sum_{ k = 2 }^n
    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
  \; \dot q^2( t )
  .
\notag
\end{align}
%
As the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
the above Lagrangian contains only the kinetic energy.
Thus, the above mechanical analogy
represents a free particle
carrying a position-dependent mass.
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved.
But in this case the Hamiltonian
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L,
  \notag
\end{equation}
coincides with the Lagrangian.
So we may set
\begin{equation}
  \mathcal L
  =
    \sum_{ k = 2 }^n
      \Gamma_k \, \lambda_k^2
      \, u_k^2 \bigl[ q(t) \bigr]
  \;
  \dot q^2(t)
  =
  \left(
    \frac C T
  \right)^2
  ,
  \label{eq:Lagrangian_const}
\end{equation}
for some positive constant $C$.
This serves
as a generalization of Eq. \eqref{eq:dudt_const}.
By integrating the square root, we get
\begin{equation}
  \int_{ 0 }^{ q(t) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  =
  C \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
where $C$ is determined by
the boundary condition at $t = T$ as
\begin{equation}
  C =
  \int_{ 0 }^{ q(T) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  .
  \label{eq:mint}
\end{equation}
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
Since we only optimized the asymptotic error, $E_A$,
in the above,
the value of $q(T)$ can be further adjusted
to minimize the total error, $E = E_R + E_A$
(cf. Sec. \ref{sec:optinitalpha}).
A complete procedure of computing the optimal schedule
is listed in Appendix \ref{sec:procedure}.

\subsubsection{\label{sec:mass_distr}
Characterization of optimal schedules}

We can characterize the optimal schedule
geometrically as follows.
First, we define a mass function as
\begin{equation}
  M(Q)
  =
  \sqrt{
    \textstyle\sum_{ k = 2 }^n \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
  }
  .
\label{eq:mass_func}
\end{equation}
and a normalized mass distribution as
\begin{equation}
  m(Q)
  =
  \frac{
    M(Q)
  }
  {
    \int_0^{ q(T) } M(Q') \, d Q'
  }
  =
  \frac{
    M(Q)
  }
  {
    C
  }
  ,
\notag
\end{equation}
such that
$\int_0^{q(T)} m(Q) \, dQ = 1$.

\begin{figure}[h]
\begin{center}
  \begin{tikzpicture}[very thick, scale=2.5, every node/.style={scale=1.5}]
    \draw[domain=0:2.5, variable=\x, dashed, gray, ultra thick]
      plot({\x}, {1.5*exp(-0.5*\x*\x)});

    \fill [gray!30!white, domain=0.7:2.2, variable=\x]
      (0.7, 0) --
      plot ({\x}, {1.5*exp(-0.5*\x*\x)})
      -- (2.2, 0) -- cycle;

    \draw[domain=0:2.2, variable=\x, ultra thick]
      plot({\x}, {1.5*exp(-0.5*\x*\x)});

    \node[] at (1.1, 0.5) {$\dfrac{t}{T}$};

    \draw[<->] (0.6, 0)  --
               node[left, inner sep=1mm]
               {$\dfrac{1}{T \, \alpha}$}
               (0.6, 1.174);

    \draw[] (0.5, 1.174) -- (0.7, 1.1740);
    \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.8);

    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
    \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
  \end{tikzpicture}
  \caption{
    \label{fig:massq}
    Geometric characterization of the optimal schedule,
    $\alpha(t)$,
    via the normalized mass distribution, $m(Q)$.
  }
\end{center}
\end{figure}

Then, from Eq. \eqref{eq:q_opt},
we find that for $Q = q(T) - q(t)$,
%
\begin{align}
  m(Q)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
\label{eq:mQ_invTa}
  \\
  \int_Q^{ q(T) }
    m(Q') \, dQ'
  &=
  \frac t T
  ,
\label{eq:intmQ_tT}
\end{align}
%
which means that each point
$\bigl(t, \alpha(t)\bigr)$
on the schedule can be mapped
to a point
$\bigl(Q, m(Q)\bigr)$
on the mass distribution,
such that the ordinate of the latter curve
gives $1/(T\alpha)$,
and the area under the curve in the domain $[Q, q(T)]$
is equal to $t/T$,
as shown in Fig. \ref{fig:massq},

This mapping allows us to classify optimal schedules
as follows.
We collect in each class optimal schedules
computed from simulations with the same function $M(Q)$
and cutoff value $q(T)$.
The former is determined by the values of
$\lambda_k$'s and $\Gamma_k$'s,
whereas the latter affects the distribution $m(Q)$
via the normalization constant $C$.

In each class, the updating magnitudes of two schedules
of lengths $T_1$ and $T_2$
at corresponding times $t_1/T_1 = t_2/T_2$
are related by $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
Further, by Eqs.
\eqref{eq:error_asym_Lagrangian}
and
\eqref{eq:Lagrangian_const},
we have
%
\begin{equation}
  E_A
  =
  \frac { C^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(Q) \, dQ
  \right)^2
  ,
\label{eq:error_asym2}
\end{equation}
%
which means the asymptotic error
is inversely proportional to the simulation length
for schedules in the same class,
i.e., we have the scaling relation
$T_1 \, E_{A1} = T_2 \, E_{A2}$.

\subsubsection{\label{sec:optinitalpha}
  Optimal initial updating magnitude
}

The above characterization allows us to determine
the optimal initial updating magnitude.
Let us now consider simulations in which
the system is initially equilibrated at the same
constant updating magnitude $a_0$.
For these simulations,
we can adjust $q(T)$
to minimize the total error.
From Eqs. \eqref{eq:error_res1}
and \eqref{eq:error_asym2},
we find that
the $q(T)$ that minimizes
the total error satisfies
\begin{equation}
  m\bigl( q(T) \bigr)
  =
  \frac{1} { T \, (a_0 / 2) }
  .
\label{eq:opt_qT}
\end{equation}
By Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
we can rewrite the condition as
\begin{equation}
  \alpha( t = 0 )
  =
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
i.e., the optimal initial updating magnitude
is always half of the equilibration value.
This factor $1/2$
is precisely the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.

Let us now consider
simulations with the same $a_0$.
Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
Eq. \eqref{eq:opt_qT} shows that
the optimal $q(T)$ increases with
the simulation length $T$,
more so if $M(Q)$ has a long tail.
This means that, if many $\lambda_k$'s are near zero,
both the optimal schedule and
normalized asymptotic error, $E_A \cdot T$,
would be sensitive to the simulation length, $T$.

\subsection{\label{sec:band-matrix}
Homogeneous updating schemes}

Many commonly-used updating schemes
can be characterized by a rigid window
centered on the current bin.
We call these schemes homogeneous,
or translationally invariant.
An example is the Gaussian updating scheme
used in metadynamics.

By the translational invariance,
homogeneous updating schemes
satisfy detailed balance,
Eq. \eqref{eq:w_detailedbalance},
for the flat distribution, $\rho_i = 1/n$ -
i.e., the updating matrix, $\mathbf w$,
is symmetric.
The updating matrices
share the same set of eigenvectors
(sines and cosines of different frequencies),
and thus are completely determined
by the eigenvalues.
Further, the multiplication of $\mathbf w$
can be reduced to a convolution with the window function,
or the updating kernel.
Thus,
there is a one-to-one correspondence between
the eigenvalues of $\mathbf w$
and the updating kernel.

\subsubsection{\label{sec:bandkernel}
Updating kernel and eigenvalues}

It is natural to first introduce homogeneous updating schemes
for a periodic variable\cite{dama2014}.
The updating matrix, $\mathbf w$,
assumes the following form
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
\notag
\end{equation}
where
$\mu_{-b}, \dots, \mu_b$ ($b \le n/2$)
are a set of numbers satisfying
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  ,
\label{eq:mu_normalization}
\end{equation}
and $\mu_l = 0$ for $|l| > b$.
We will further impose the symmetry
\begin{equation}
  \mu_i = \mu_{-i}
  .
\label{eq:mu_symm}
\end{equation}
If $b = n/2$, then $\mu_{-b}$ is removed
from the sum in Eq. \eqref{eq:mu_normalization}
to avoid double counting.
These numbers constitute an \emph{updating kernel}.
In this way, the matrix, $\mathbf w$,
being symmetric with respect to the indices $i$ and $j$,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution.

To find the eigenvectors,
we define, for a periodic variable, $\phi_i$,
the out-of-boundary values by
\begin{equation}
  \phi_i = \phi_{i \pm n},
\label{eq:phi_pbc}
\end{equation}
such that $i \pm n$ lies in between $1$ and $n$.
Then, the multiplication of the matrix, $\mathbf w$,
is equivalent to a convolution with the kernel:
\begin{equation}
  \sum_{ j = 1 }^n
    w_{ij} \, \phi_j
  =
  \sum_{ j = 1 - n }^{ 2 \, n }
    \mu_{i - j} \, \phi_j
  =
  \sum_{ l = -b }^{ b }
    \mu_l \, \phi_{ i - l}
  .
\label{eq:wmul_to_convol}
\end{equation}
As a result, the characteristic equation,
$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
can be solved by discrete Fourier transform.
We may construct a set of orthonormal eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
compatible with the periodicity, Eq. \eqref{eq:phi_pbc},
\begin{equation}
  \phi^{(k)}_j
  =
  \phi_{kj}
  =
  \frac 1 n
  \exp
  \left[
    \ii
    \frac{ (k - 1) \, j \, 2 \, \pi }
         {            n             }
  \right]
  ,
  \notag
\end{equation}
where $\ii^2 = -1$, and
the eigenvalues are
the Fourier or cosine transform of the updating kernel:
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ l = 1 }^b
  \mu_l
  \cos
  \frac{ (k - 1) \, l \, 2 \, \pi }
       {            n             }
  .
  \label{eq:wband_eigenvalue_pbc}
\end{equation}
Note, however, the two-fold degeneracy as
$\lambda_{n - k + 2} = \lambda_k$.

To adapt the above formulae to a non-periodic variable,
we use the reflective boundary condition\cite{bussi2006},
and change the form of the updating matrix to
\begin{equation}
  w_{ij}
  =
  \mu_{ i - j }
  +
  \mu_{ i + j - 1 }
  +
  \mu_{ i + j - 2 n - 1 }
  ,
  \label{eq:w_band_refl}
\end{equation}
where the last two terms on the right hand side,
representing two reflective mirrors at
$j = 1/2$ and $j = n + 1/2$,
are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
to the equilibrium distribution\cite{bussi2006},
such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
are satisfied.
Again the updating kernel should satisfy
Eqs. \eqref{eq:mu_normalization}
and \eqref{eq:mu_symm},
but
the cutoff $b$ only needs to be less than $n$.

Parallel to Eq. \eqref{eq:phi_pbc},
we define the out-of-boundary values
of $\phi_i$ as
\begin{equation}
  \phi_i
  =
  \begin{dcases}
    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
  \end{dcases}
\label{eq:phi_refl}
\end{equation}
such that the matrix multiplication is still equivalent to
a convolution as Eq. \eqref{eq:wmul_to_convol}.
We note that Eq. \eqref{eq:phi_refl}
defines a periodic variable of period $2 \, n$
with a reflective symmetry around $i = 1/2$.
Accordingly, the eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
must share the same periodicity,
and be even functions about the axis $i = 1/2$:
\begin{equation}
  \phi^{(k)}_i
  =
  \phi_{k i}
  =
  \frac{ \sqrt{ 2 - \delta_{k, 1} } }
       {             n              }
  \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, 2 \, \pi}
            {                    2 \, n                           }
  ,
\notag
\end{equation}
The eigenvalues can be borrowed from
Eq. \eqref{eq:wband_eigenvalue_pbc},
with the period $n$ doubled:
\begin{align}
  \lambda_k
  &=
  \mu_0
  +
  2
  \sum_{l = 1}^b
    \mu_l
    \cos \frac{(k - 1)  \, l \, 2 \, \pi}{ 2 \, n }
  .
\label{eq:wband_eigenvalue_refl}
\end{align}
We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
and \eqref{eq:wband_eigenvalue_pbc} as
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2
  \sum_{ l = 1 }^b
    \mu_l \,
    \cos \frac{ (k - 1) \, l \, g \, \pi }
              {            n             }
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
where the degeneracy $g$ is $2$
for the periodic case,
or $1$ otherwise.
In both cases,
the eigenvalues, $\lambda_k$'s,
are the cosine transform of
the updating kernel.
For the updating scheme to be stable,
the cosine transform of the kernel
must be nonnegative.
If this is not so,
we can modify the kernel
using the technique described
in Appendix \ref{sec:stabilize_wband}.
We discuss two special cases below.

\subsubsection{\label{sec:nnscheme}
Triple-bin updating scheme}

If $\mu_2 = \dots = \mu_b = 0$,
the matrix $\mathbf w$ is tridiagonal,
and the eigenvalues are
\begin{equation}
  \lambda_k
  =
  1 -
  4 \, \mu_1 \sin^2
  \frac{ (k - 1) \, g \, \pi }
       {       2 \, n        }
  .
\notag
\end{equation}
So this updating scheme is stable if
$\min \lambda_k \ge 0$,
or
\begin{equation}
  \mu_1 \le
  \begin{dcases}
    \frac 1 4
    & \mathrm{periodic \; and \;} n \mathrm{ \; even,}
    \\
    \frac 1 4
    \sec^2
    \left( \frac{  \pi   }
                { 2 \, n }
    \right)
    & \mathrm{otherwise}.
  \end{dcases}
\notag
\end{equation}

\subsubsection{Gaussian updating scheme}

We define a Gaussian updating scheme
as an updating scheme with a
Gaussian-shaped kernel.
This updating scheme is commonly
used in metadynamics.
Below we show its stability
in the continuous limit.

For $n \gg 1$,
we define
$x = l \, \Delta x$,
and
$\mu_l = \mu(x) \, \Delta x$,
where
$\Delta x = g \, \pi/n$
is the bin size.
By assuming the largest $b \approx n/g$,
we can approximate Eq. \eqref{eq:wband_eigenvalue}
as an integral:
\begin{equation}
  \lambda_k
  =
  2 \int_0^\pi
    \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
\notag
\end{equation}
with the normalization
$
  1 = 2 \int_0^\pi \mu(x) \, dx.
$

For the Gaussian updating kernel,
if the width $\sigma$ satisfies $\Delta x \ll \sigma \ll 1$,
we can extend
the upper limit of the integrals
to infinity, and
\begin{equation}
  \mu(x)
  =
  \frac{            1            }
       { \sqrt{ 2 \pi \sigma^2 } }
  \exp\left(
        -\frac{       x^2     }
              { 2 \, \sigma^2 }
      \right)
  .
\notag
\end{equation}
If follows that all eigenvalues are positive:
\begin{align}
  \lambda_k
  &=
  \exp\left[
        -\frac{ (k - 1)^2 \, \sigma^2 }
              {           2           }
      \right]
  .
\notag
\end{align}

\subsubsection{\label{sec:invert_wband}
Inversion}

Since the eigenvalues are related to the updating kernel
by a cosine transform,
Eq. \eqref{eq:wband_eigenvalue},
the relation can be readily inverted.
In the periodic case,
we have
\begin{equation}
  \mu_l
  =
  \frac 1 n
  \sum_{ k = 1 }^n
  \lambda_k
  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
            {            n             }
  .
\label{eq:mu_from_lambda_pbc}
\end{equation}

In the non-periodic case,
the formula is
\begin{align}
  \mu_l
  =
  \frac 1 { 2 \, n }
  \sum_{ k = 1 }^{ 2 \, n }
    \lambda_{ k } \,
    \cos \frac{ (k - 1) \, l \, \pi }
              {            n        }
  ,
\label{eq:mu_from_lambda_refl}
\end{align}
where
we have defined
$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
for $k = n + 2, \dots, 2 \, n$,
as well as
\begin{align}
  \lambda_{ n + 1 }
  =
  (-1)^{ n - 1 }
  \lambda_1
  +
  2 \, \sum_{ k = 2 }^{ n }
      (-1)^{n - k} \, \lambda_k
  ,
\label{eq:lambdan}
\end{align}
to satisfy the constraint $\mu_n = 0$.

\subsection{\label{sec:cmpschemes}
Comparison of updating schemes}

Now that we can compute the optimal schedule
for a given updating scheme,
we will turn to the problem of finding
optimal updating schemes below.

\subsubsection{\label{sec:optWL}
Asymptotic optimality of the single-bin scheme}

We will first show
that the single-bin scheme is asymptotically optimal.
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence the same $\Gamma_k$'s,
but with different positive eigenvalues,
$\lambda_k$'s.
Using the Cauchy-Schwarz inequality, we have,
for any set of nonnegative numbers, $c_k$'s,
\begin{align}
&
\left(
  \int_0^T dt
    \sum_{k = 2}^n
      \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr)
\right)
\left(
  \int_0^T dt
    \sum_{k = 2}^n
      \Gamma_k \, c_k^2
\right)
\notag
\\
&
\qquad \qquad
\ge
\left(
  \int_0^T dt
    \sum_{k = 2}^n
      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t) \bigr)
\right)^2
\notag
\\
&
\qquad \qquad
=
\left(
  \sum_{k = 2}^n \Gamma_k \, c_k
    \left[
      1 - e^{ -\lambda_k \, q(T) }
    \right]
\right)^2.
\label{eq:CSineq}
\end{align}
Note that the last expression of Eq. \eqref{eq:CSineq}
depends on the curve, $q(t)$ ($0 < t < T$),
only through the endpoint value, $q(T)$,
which is fixed in the variational process.
Thus, the inequality sets a lower bound
for the asymptotic error $\Err$
given in Eq. \eqref{eq:error_asym1}.
The equality is achieved
if $\dot u_k\left( q(t) \right) = c_k$
for all $k = 2, \dots, n$ at any $t$,
up to a multiplicative factor.
Solving these equations yields
\begin{equation}
  \alpha(t) = \frac{              1             }
                   { \lambda_k \, (t + t_{k0} ) },
\label{eq:alpha_invtk}
\end{equation}
with
\begin{equation}
  t_{k0} = \frac{             T            }
                { e^{ \lambda_k q(T) } - 1 }.
\label{eq:tk0}
\end{equation}
Such a solution is possible only if
all nontrivial eigenvalues are the same
\begin{equation}
  \lambda_2 = \dots = \lambda_n.
  \notag
\end{equation}
Then, by Eq. \eqref{eq:tk0},
all $t_{k0}$'s also share the same value,
$t_0$,
such that
$u_k = (t + t_0) / (T + t_0)$,
and
$c_k = 1/(T + t_0)$
for $k \ge 2$.
The asymptotic error is then
\begin{align}
  E_A
  =
  \frac{       T     }
       { (T + t_0)^2 }
  \sum_{ k = 2 }^n
    \Gamma_k
  .
\label{eq:error_asym_singlebin}
\end{align}
If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
Eq. \eqref{eq:alpha_invtk}
recovers Eq. \eqref{eq:alpha_invt1}.

Since the first eigenmode represents
a uniform shift of the bias potential,
we can freely set $\lambda_1$ to be the same as $\lambda_2$
without changing the nature of the updating scheme.
Note, an updating matrix
with identical eigenvalues
is a multiple of the identity matrix.
Thus,
the corresponding updating scheme,
the single-bin scheme, is most efficient
in terms of asymptotic convergence.
In Appendix \ref{sec:upperbound},
we will show that the opposite limit of
a set of well-separated eigenvalues
will provide an upper bound for
the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.

\subsubsection{\label{sec:optscheme}
Bandpass updating schemes}

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that ignore short-wave eigenmodes,
which are typically local noise.
Mathematically, these updating schemes will have
zero eigenvalues for the ignored modes,
and they are useful if we can assume that
the ignored modes are absent from the PMF.
For example,
by assuming
that the PMF is sufficiently smooth
such that the eigenmodes can be limited to $k \le K$,
we can construct an updating scheme with
$$
\lambda_1 = \cdots = \lambda_K = 1,
\mathrm{\; and \;}
\lambda_{K+1} = \cdots = \lambda_n = 0.
$$
Then,
we can change the upper limit of the sum in
Eq. \eqref{eq:error_asym_singlebin} to $K$:
\begin{equation}
  E_A
  =
  \frac {       T     }
        { (T + t_0)^2 }
  \sum_{ k = 2 }^K
    \Gamma_k.
\label{eq:error_asym_sinc}
\end{equation}
The optimal schedule for these types of schemes
is given by the inverse-time scheme (Eq. \eqref{eq:alpha_invt1}).
Below, we construct these schemes for
the homogeneous case.
The definition of the cutoff $K$ will be modified
for mathematical convenience.

For a periodic variable, we demand
$$
\begin{aligned}
&
\lambda_1 = \cdots = \lambda_{K+1}
= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
\\
&
\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
\end{aligned}
$$
for $K \ge 1$.
Then by using
Eq. \eqref{eq:mu_from_lambda_pbc},
we get
\begin{equation}
  \mu_l
  =
  \frac{
    \sin
    \dfrac{ (2 K + 1) \, l \, \pi }
         {              n        }
  }
  {
    n \, \sin \dfrac{ l \, \pi } { n }
  }
  .
\notag
\end{equation}

For a non-periodic variable, we demand
$$
\lambda_1 = \cdots = \lambda_{K+1} = 1,
\qquad
\lambda_{K+2} = \cdots = \lambda_n = 0.
$$
Then Eq. \eqref{eq:lambdan} gives
$\lambda_{n+1} = (-1)^{n+K-1}$,
and from Eq. \eqref{eq:mu_from_lambda_refl},
we have
\begin{equation}
  \mu_l
  =
  \frac{1}{2 \, n}
  \left[
    (-1)^{n+K+l-1}
    +
    \frac{
      \sin
      \dfrac{ (2 K + 1) \, l \, \pi }
           {         2 \, n        }
    }
    {
      \sin \dfrac{ l \, \pi } { 2 \, n }
    }
  \right]
  .
\label{eq:mu_sinc_refl}
\end{equation}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
  }
  \caption{
    \label{fig:sinc}
    Updating kernels of the bandpass updating schemes
    compared to a Gaussian with a similar width.
    The lines are a guide to the eyes.
    Inset. Updating matrix associated with the non-periodic case.
  }
\end{center}
\end{figure}

In Fig. \ref{fig:sinc},
we give examples of the updating kernels
of the bandpass updating schemes,
compared to a Gaussian
with a similar width, $\sigma$,
\begin{equation}
  \sigma
  \approx
  \frac
  {
    n
  }
  {
    \sqrt{ 2 \, \pi } \, g \, K
  }
  .
\label{eq:sigma_equiv}
\end{equation}
The kernel in the non-periodic case is sawtooth
because of the term $(-1)^{n+K+l-1}$
in Eq. \eqref{eq:mu_sinc_refl}.
However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
the ruggedness disappears
in the updating matrix by Eq. \eqref{eq:w_band_refl},
as the above oscillatory term for $l = i - j$
is canceled by the terms associated with $l = i + j - 1$
or $l = i + j - 2 \, n - 1$
with opposite sign.

\subsubsection{\label{sec:finlencorr}
Finite-length correction}

In the above section, we have focused on
the asymptotic error.
By including the residual error
and minimizing the total error,
we get the optimal $t_0$,
which is, according to Eq. \eqref{eq:half_alpha0},
\begin{equation}
  t_0
  =
  \frac{  2  }
       { a_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
From Eqs. \eqref{eq:error_split}, the total error is
\eqref{eq:error_res1},
\eqref{eq:tk0},
and
\eqref{eq:error_asym_sinc},
\begin{equation}
  E
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
This error decreases inversely with increasing simulation length,
as one would expect from an equilibrium FDS simulation.

\section{\label{sec:results}
Numerical results}

We tested the above theory on a model system
described in Sec. \ref{sec:results_system}.
We verify the method of computing errors
in Sec. \ref{sec:results_invt}.
and study the optimal schedule from
Eq. \eqref{eq:q_opt}
in Sec. \ref{sec:results_optschedule}.
Finally, we compare the Gaussian and
bandpass updating schemes %of different widths
in Sec. \ref{sec:results_cmpschemes}.

\subsection{\label{sec:results_system}
Test system and simulation protocol}

Our test system is a one-dimensional model system, in which
$i$ represents a non-periodic variable,
and $n = 100$.
Both the target and intrinsic distributions
are assumed to be flat:
$p^*_i = p_i = 1/n$,
along with the distribution $\rho_i$ in Eqs.
\eqref{eq:vbar_def} and \eqref{eq:error_sum}.
Initially,
we set the bias potential to zero,
$V_i \equiv 0$,
and equilibrate the system for $10^7$ steps
under a constant updating magnitude,
$a_0 = 10^{-4}$.
Although the bias potential is correct initially,
it would deviate from this value
at the end of the equilibration.
Our aim is to see how well
the bias potential is restored
at the end of the testing schedule.
After the equilibration,
we reset the origin of the time, $t$, to zero,
and start the testing schedule
for $T = 10^8$ steps.
We repeat the above process $1000$ times (unless specified otherwise)
and average the results.

For the Gaussian updating scheme,
the kernel was truncated at
$b = \min\{10 \, \sigma, n - 1\}$
and numerically stabilized
using the technique
in Appendix \ref{sec:stabilize_wband}.

We tested two sampling processes.
Both are based on
a Metropolis MC algorithm\cite{
  metropolis1953, newman, frenkel,
  landau_binder}:
in each step, a new bin index $j$ is proposed
and then accepted with probability
$
A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
$
However,
in the first or global sampling process,
we choose the destination $j$
uniformly out of the $n$ bins.
In the second or local sampling process,
$j$ is either $i - 1$ or $i + 1$
(adjusted by the boundary condition)
with equal probability $1/2$.
Asymptotically,
the above global and local sampling processes
emulate the perfect and one-step processes
discussed in Appendix \ref{sec:Gamma},
respectively,
and we will use
Eqs. \eqref{eq:Gamma_perfect}
and \eqref{eq:Gamma_onestep}
for the respective $\Gamma_k$'s
in the theoretical predictions.

\subsection{\label{sec:results_invt}
Modified inverse-time schedule}

Using the model system above, we tested our analysis of computing errors
on the modified inverse-time schedule,
\begin{equation}
\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
\label{eq:alpha_invtlambda}
\end{equation}
where $\lambda$ is a free parameter,
and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
The error of this schedule
can be computed analytically
(cf. Appendix \ref{sec:invt_schedule}),
and tested against simulation.

We performed adaptive FDS simulations
using the single-bin updating scheme
and the triple-bin (bandpass) updating scheme with $\mu_1 = 0.24$.
As shown in Fig. \ref{fig:errinvt},
the errors from the simulations
agreed well those from the theory.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/errinvt.pdf}
  }
  \caption{
    \label{fig:errinvt}
    Error $E$ versus the parameter $\lambda$
    for the modified inverse-time schedule,
    Eq. \eqref{eq:alpha_invtlambda}.
    The points are results from averaging over $1000$ independent runs;
    the curves are theoretical predictions.
    For the single-bin (WL) updating scheme,
    the minimal should occur at $\lambda = 1$,
    while it is not necessarily the case for a general updating scheme.
  }
\end{center}
\end{figure}

For the single-bin updating scheme,
the optimal $\lambda$ occurred around $1.0$
for both global and local sampling processes.
The sampling process affected the error function
only by a multiplicative constant,
in agreement with the discussion in
Appendix \ref{sec:invt_singlebin}.
For the triple-bin updating scheme,
the optimal $\lambda$ was less than $1.0$
for the global sampling process,
although it remained around $1.0$
for the local sampling process,
as discussed in Appendix \ref{sec:invt_nn}.
Thus, the inverse-time formula,
Eq. \eqref{eq:alpha_invt1},
is optimal for the single-bin updating scheme,
but not necessarily so for a general, multi-bin updating scheme.

\subsection{\label{sec:results_optschedule}
Optimal schedule}

To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
we used the Gaussian (with $\sigma = 10$) updating scheme
as an example.
In Table \ref{tab:error_Gaussian},
we show that the optimal schedule
produced smaller errors than
the inverse-time schedule.
However, the optimal schedule
decayed more slowly than
the inverse-time formula, Eq. \eqref{eq:alpha_invt1},
and depended on the sampling process,
as shown in Fig. \ref{fig:optacmp}.
The impeded decay of the optimal schedule
was necessary in order to
damp out the short-wavelength eigenmodes
(i.e., local noise),
as shown in Fig. \ref{fig:xerr}.
If, however, we could ignore these local eigenmodes
(by explicitly filtering
them from the biased potential),
we could modify the optimal schedule
(Eq. \eqref{eq:q_opt})
and limit the sum
to the first few eigenmodes.
The resulting schedule
would become similar to the inverse-time schedule,
Eq. \eqref{eq:alpha_invt1},
as shown in Fig. \ref{fig:optacmp}.

\begin{table}[h]
  \caption{\label{tab:error_Gaussian}
    Error of the Gaussian updating scheme of $\sigma = 10$.
    $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.2}
  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
    \hline
      Schedule
    &
      Inverse-time, \newline
      Eq. \eqref{eq:alpha_invt1}
    &
      Modified inverse-time,
      Eq.~\eqref{eq:alpha_invtlambda}
      with optimized $\lambda$
    &
      Optimal, \newline
      Eq. \eqref{eq:q_opt}
    \\
    \hline
    Global
    &
    $4.36(9) \times 10^{-6}$
    &
    $7.1(1) \times 10^{-7}$
    &
    $2.6(1) \times 10^{-7}$
    \\
    \hline
    Local
    &
    $3.83(7) \times 10^{-4}$
    &
    $1.72(4) \times 10^{-4}$
    &
    $9.1(2) \times 10^{-5}$
    \\
    \hline
  \end{tabular}
\end{table}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/optacmp.pdf}
  }
  \caption{
    \label{fig:optacmp}
    Optimal schedules from Eq. \eqref{eq:q_opt}
    for the Gaussian updating scheme
    with $\sigma = 10$.
    The mode-limited schedules were produced by
    limiting the sum in Eq. \eqref{eq:q_opt}
    to the first four terms.
    They became similar to the inverse-time schedule,
    Eq. \eqref{eq:alpha_invt1}.
  }
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/xerr.pdf}
  }
  \caption{
    \label{fig:xerr}
    Predicted error components for the Gaussian updating scheme
    with $\sigma = 10$
    from the optimal schedule
    and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
    The latter schedule failed to damp out the errors
    from the long-wavelength (local) eigenmodes
    with $k \ge 5$.
    For longer simulations,
    the error components from the optimal schedule
    would roughly maintain the relatively flat profile,
    but meet the initial error curve
    at a lower level (not shown),
    where $\lambda_k \approx 1/q(T)$
    (cf. Appendix \ref{sec:upperbound}).
  }
\end{center}
\end{figure}

To verify Eq. \eqref{eq:half_alpha0},
we computed the optimal schedules
for different values of $q(T)$,
and plotted the final error
against the initial updating magnitude, $\alpha(0)$.
As shown in Fig. \ref{fig:erriascan},
the minimal error was reached at
$\alpha(0) = a_0/2 = 5 \times 10^{-5}$.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/erriascan.pdf}
  }
  \caption{
    \label{fig:erriascan}
    Relative error
    versus the initial updating magnitude, $\alpha(0)$,
    for optimal schedules with different $q(T)$.
    Assuming the Gaussian updating scheme
    with $\sigma = 10$,
    and $a_0 = 10^{-4}$.
    The points were obtained from averaging results
    from over 10000 independent simulations;
    the curves are theoretical predictions.
    The minimal should always occur at
    $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
    as predicted by Eq. \eqref{eq:half_alpha0}.
  }
\end{center}
\end{figure}

\subsection{\label{sec:results_cmpschemes}
Optimal updating schemes}

For a finite-length simulation,
the optimal updating scheme
depends on the simulation length, $T$,
and other parameters.
As an example,
we compare Gaussian updating schemes
with different widths, $\sigma$.
As shown in Fig. \ref{fig:errsigscan}(a)
that the minimal error was achieved
by the widest window %for the shorter simulation of $T = 10^8$,
for the global sampling process,
although
the minimum at $\sigma = 0$ became more favorable
as the simulation lengthens.
However,
for the local sampling process,
the minimum stayed at zero width,
as shown in Fig. \ref{fig:errsigscan}(b).

In contrast,
the normalized errors of the bandpass updating schemes
were insensitive to the simulation length,
in agreement with Eq. \eqref{eq:error_sinc}.
The bandpass updating schemes
yielded smaller errors than
the corresponding Gaussian ones,
although the differences tended to
be small for shorter simulations
with the local sampling process.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsigscan.pdf}
  }
  \caption{
    \label{fig:errsigscan}
    Normalized error, $(T + t_0) \, E$,
    versus the width, $\sigma$,
    for the Gaussian
    and the bandpass updating schemes,
    with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
    In the latter case,
    the equivalent width
    $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
    The points were obtained from averaging results
    from 1000 independent simulations;
    the curves are theoretical predictions.
  }
\end{center}
\end{figure}

\section{\label{sec:conclusion}
Conclusions and Discussions}

In conclusion,
we have proposed a method of computing
the optimal schedule of the updating magnitude
for a general class of free energy calculation methods,
which we refer to as adaptive flat-distribution sampling (FDS) methods,
that include the Wang-Landau (WL) algorithm and metadynamics.
Adaptive FDS methods
incrementally update a bias potential
along a collective variable
to offset the potential of mean force (PMF)
and thereby sample a flat distribution.
The optimal schedule delivers the fastest convergence
of the bias potential,
and thus helps minimize the effort
of computing the PMF.

The optimal schedule depends on the updating window function,
which is limited to a single-bin in the WL case,
but can be extended to several neighboring bins
taking a Gaussian shape in the metadynamics case.
More generally,
each adaptive FDS method is associated with
an updating scheme
that can be mathematically characterized
by an updating matrix.
Each column of the matrix gives the normalized
window function for the bin corresponding to the column.
Our method computes the optimal schedule from
the eigenvalues of the updating matrix,
and the integrals of the autocorrelation functions
of the eigenmodes.
For the single-bin (WL) scheme,
the optimal schedule
is given by the inverse-time formula,
Eq. \eqref{eq:alpha_invt1}.
For a general multiple-bin scheme,
including the Gaussian one used in metadynamics,
the optimal schedule is not given by a simple closed form,
but is implicitly given by an equation of motion
of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
In this case,
the optimal schedule and error
can be sensitive to the simulation length.
The initial updating magnitude is always
half of the previous equilibrium value.

We have also shown that
the single-bin scheme is optimal
in the long time limit
without a priori assumptions
about the smoothness of the PMF.
However, if certain eigenmodes associated with local noise
are expected to vanish in the PMF,
we can construct a
class of bandpass updating schemes
that are also optimal for asymptotic convergence.
The optimal schedule of these bandpass schemes
is always given by an inverse-time formula.

Although the bandpass updating schemes
appeared to outperform similar Gaussian schemes
in our model study,
they are global in nature
and hence more expensive.
However, the performance gain
is relatively small for short simulations
with a local sampling process such as MD.

In this analysis, we have ignored the \emph{systematic} error\cite{zhou2005}
caused by the updates in adaptive FDS methods.
This is a valid assumption in the asymptotic regime,
where the random error is much greater than the systematic one.
This allows us to assume the sampling process
to be a quasi-equilibrium one\cite{barducci2008, dama2014},
and to represent the autocorrelation function
of each eigenmode by a single number, $\Gamma_k$.
Although this assumption appeared to work well
on our test system for modeling long simulations,
they may break down for relatively short simulations
of glassy systems.

\section{Acknowledgments}

We thank Dr. Y. Mei, J. Wang,
O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
Dr. C. Myers, and Dr. O. Samoylova
for helpful discussions.  We gratefully acknowledge the Robert A. Welch Foundation (H-0037), the National Science Foundation (CHE-1152876) and the National Institutes of Health (GM-037657) for partial support of this work.  This research used computing resources of the National Science Foundation XSEDE grid.

\appendix

\section{\label{sec:equilerr}
Inverse-time schedule as a result of increasing sample size
in a runtime average
}

We can understand the optimality of the inverse-time schedule
as a result of the stabilization of a runtime average
with respect to the increase of sample size\cite{marsili2006, barducci2008}.
Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
with a dynamically growing size.
The runtime average
\begin{equation*}
  {\bar X}^{(N)}
  =
  \frac{ X^{(1)} + \cdots + X^{(N)} } { N }
  ,
\end{equation*}
can be written as a recurrence relation regarding
the addition of the last data point.
\begin{align}
  {\bar X}^{(N)}
  &=
  {\bar X}^{(N-1)} + \frac{ X^{(N)} - {\bar X}^{(N-1)} } { N }
  .
  \label{eq:av_recur}
\end{align}
If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
as the independent correction to the previous average,
${\bar X}^{(N-1)}$, proposed by the latest data point, $X^{(N)}$,
then this correction is absorbed into the average with
a diminishing weight $1/N$ as the sample size, $N$, increases.

Let us now consider a long equilibrium FDS simulation
that consists of blocks of sub-simulations
of $L$ steps.
If $X^{(N)}$ is
the bias potential, $V_i^{(N)}$, computed
from the $N$th sub-simulation,
then ${\bar X}^{(N)}$ stands for the average bias potential,
${\bar V}_i^{(N)}$,
from the first $N$ sub-simulations,
which is the best estimate at the end of
the first $N$ sub-simulations.
If we wish to always use this best estimate, ${\bar V}_i^{(N-1)}$,
from the previous $N-1$ sub-simulations
as the bias potential for the $N$th sub-simulations,
then this estimate can be updated according to Eq. \eqref{eq:av_recur}.
Assuming that the normalized histogram, $H_i^{(N)}$,
from the $N$th sub-simulations is sufficiently precise,
Eq. \eqref{eq:vcorr_equil} shows that the independent
correction from the $N$th sub-simulation is
\begin{equation*}
  \ln \frac{ H_i^{(N)} } { p_i }
  \approx
  \frac{ H_i^{(N)} } { p_i } - 1
  ,
\end{equation*}
where, the $-1$ on the right-hand side
can be ignored as a uniform shift that
does not affect the distribution.
Thus, we may rewrite Eq. \eqref{eq:av_recur} as
\begin{align}
  {\bar V}_i^{(N)}
  &=
  {\bar V}_i^{(N-1)}
  +
  \frac{1}{N} \frac{ H_i^{(N)} }{ p_i }
  \notag
  \\
  &=
  {\bar V}_i^{(N-1)}
  +
  \sum_{l=1}^L \frac{1}{N \, L} \frac{ h_i^{(N)}(l) }{ p_i }
  ,
  \label{eq:vbar_update}
\end{align}
where $h_i^{(N)}(l)$ stands for the instantaneous histogram
at step $l$ of sub-simulation $N$,
and it is related to the average $H_i^{(N)}$ as
$H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
Thus, Eq. \eqref{eq:vbar_update} shows that the best estimate
or the runtime average of the bias potential can be computed
through a step-by-step incremental updating scheme that is equivalent to
the WL algorithm, Eq. \eqref{eq:wl_update},
with the magnitude being
\begin{equation}
  \alpha(t) = 1/(NL)
  .
  \label{eq:alpha_block}
\end{equation}
In the long time limit,
we may approximate the simulation time
by the product of the number of sub-simulations
and the number of steps in each sub-simulation, or $NL$,
and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.

The above demonstration has several practical implications.
First,
since the bias potential is often sufficiently precise
in late stages of a WL simulation,
some practitioners would then fix
the bias potential and switch
to the equilibrium FDS method
to prevent possible future efficiency loss
due to the adaptive updates.
However, the above demonstration shows that
this practice becomes unnecessary
with the inverse-time schedule.
That is,
the WL algorithm with the inverse-time schedule
produces a bias potential that is similar in precision to that from
the ideal equilibrium simulation for long times.

Second,
an alternative way of improving the precision of the bias potential
from a simulation using the WL algorithm or metadynamics
with a fixed magnitude
is to time average the bias potential\cite{zhou2005}.
This practice becomes redundant with the inverse-time schedule
because the averaging of the bias potential
has been implicitly performed by the schedule.
The inverse-time schedule approach is preferred, however,
because with a decreasing updating magnitude,
it minimizes the systematic error caused by
the on-the-fly updating scheme\cite{zhou2005},
which is ignored in this analysis.

Third,
it is possible to reformulate the WL algorithm
with the inverse-time schedule
such that the bias potential is computed from
a weighted runtime average of the histogram\cite{marsili2006, barducci2008},
which is analogous to the adaptive bias force method\cite{darve2001, *darve2008}.

\section{\label{sec:wltoinvt}
Inverse-time schedule as the optimal limit of the WL algorithm}

We may understand the inverse-time schedule %, Eq. \eqref{eq:},
as an optimal limit of the stage-wise WL updating scheme.
Our calculation below will focus on the one-variable problem,
as in Sec. \ref{sec:onevar}.

To begin, we split the entire simulation into $S$ stages,
each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
We denote the number of steps in stage $s$ as $T_s$, and
$\sum_{s=1}^S T_s = T$.
We assume that
the magnitude follows geometrically,
\begin{equation}
  \alpha_s = \alpha_0 \, r^s
  ,
  \label{eq:alphas}
\end{equation}
with $0 < r < 1$.
We wish to find the best distribution of the simulation time
across the $S$ stages
as well as
the optimal number of stages and $r$.

If we denote the error at the end of stage $s$ as $E_s$,
then from Eq. \eqref{eq:x2t_average}, we have
$q(t) = \alpha_s \, t$, and
\begin{align}
  E_s
  &=
  E_{s - 1} \, e^{-2 \, \alpha_s \, T_s }
  \notag\\
  &\phantom{=}
  +
  \int_0^{T_s} \int_0^{T_s}
  \alpha_s^2 \,
  e^{ -\alpha_s \, (T_s - t) }
  e^{ -\alpha_s \, (T_s - t') }
  \kappa( t - t' ) \, dt \, dt'
  \notag
  \\
  &=
  \left(
    E_{s - 1} - \frac{ \Gamma \, \alpha_s } { 2 }
  \right)
  e^{ - 2 \, \alpha_s \, T_s }
  +
  \frac{ \Gamma \, \alpha_s } { 2 }
  .
  \label{eq:errorstage}
\end{align}
where we have assumed
$\kappa(t) = \Gamma \, \delta(t)$
in the second step.
Our aim is to minimize the error
of the final stage, $E_S$.
To do so, let us consider moving simulation by $\delta T$ time from
stage $s+1$ to stage $s$,
then $E_{s+1}$ is affected as
\begin{align*}
  \delta E_{s+1}
  =
\left[
  2 \, \alpha_{s+1}
  \left(
    E_s - \tfrac{ \Gamma \, \alpha_{s+1} } { 2 }
  \right)
  +
  \frac{ \partial E_s } { \partial T_s }
\right]
e^{-2 \, \alpha_{s+1} \, T_{s+1} }
\,
\delta T
.
\end{align*}
Since this variation does not directly affect the error of later stages,
the above expression must be zero to minimize the final error, i.e,
\begin{align*}
  2 \, \alpha_{s+1}
  \left(
    E_s - \frac{ \Gamma \, \alpha_{s+1} } { 2 }
  \right)
  &=
  -\frac{ \partial E_s } { \partial T_s }
  =
  2 \, \alpha_s \, \left(E_s - \frac{ \Gamma \, \alpha_s } { 2 }  \right)
  ,
\end{align*}
where we have used Eq. \eqref{eq:errorstage}
in the second step.
Thus,
\begin{equation}
  E_s
  =
  \frac{ \Gamma } { 2 } \, ( \alpha_s + \alpha_{s+1} )
  =
  \frac{ \Gamma } { 2 } \, \alpha_s ( 1 + r )
  .
  \label{eq:Es}
\end{equation}
Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
in Eq. \eqref{eq:errorstage}, we get
\begin{equation}
  e^{-\alpha_s \, T_s } = r
  .
  \label{eq:exp_r}
\end{equation}

To compute the optimal schedule,
we map the simulation time $t$ to the stage number $s$ as
\begin{align*}
  t
  = \sum_{s' = 1}^s T_{s'}
  = \frac{ 1 - r^s } { 1 - r } \, T_s
  = \frac{ 1 } { \gamma } \,
  \left(
    \frac{1}{\alpha_s}
    -
    \frac{1}{\alpha_0}
  \right)
  ,
\end{align*}
where
$\gamma = ( 1 - r ) / \ln(1/r)$,
and we have used Eq. \eqref{eq:exp_r} in the last step.
Thus,
%
\begin{equation}
  \alpha_s =
  \frac{ 1 }
  { \gamma \, t + \alpha_0^{-1} }
  ,
  \label{eq:alphas_t}
\end{equation}
%
and the final error is given by
%
\begin{align}
  E_S
  =
  \frac{\Gamma}{2}
  \frac{ 1 + r }
  { \gamma \, T + \alpha_0^{-1} }
  ,
  \notag
\end{align}
which is minimized as $r \to 1$
for a large $T$
(at this point $\gamma \to 1$).
If we identify the parameter $\alpha_0$ as the initial updating magnitude $\alpha(t = 0)$,
then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
and the optimal $t_0 = \alpha_0^{-1}$.
Thus, in this limit,
the schedule given by Eq. \eqref{eq:alphas_t}
is identical to that from Eq. \eqref{eq:alpha_invt1}.

\section{\label{sec:Gamma}
Integrals of the autocorrelation functions}

Here, we evaluate the integrals of
the autocorrelation functions,
$\Gamma_k$'s,
in two special cases
in Secs. \ref{sec:Gamma_perfect}
and \ref{sec:Gamma_onestep}
in the asymptotic limit,
and describe a general method of measuring them
in Sec. \ref{sec:Gamma_measure}.

\subsection{\label{sec:Gamma_perfect}
Perfect sampling}

First, if the sampling is perfect,
the bin index $i$ is an independent random variable
at each time step $t$.
Further, in the asymptotic limit,
$\langle h_i(t) \rangle = p_i$,
we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
%
\begin{equation*}
  \xi_i(t)
  =
  \frac{ h_i(t) } { p_i }
  -
  \sum_{ r = 1 }^n
    \rho_r \frac{ h_r(t) } { p_r }
  ,
\end{equation*}
%
and
%
\begin{align}
  \left\langle
    \xi_i(t) \, \xi_j(t')
  \right\rangle
  &=
  \left(
    \frac{ \delta_{ij} } { p_i }
    -
    \frac{ \rho_i } { p_i }
    -
    \frac{ \rho_j } { p_j }
    +
    \sum_{r = 1}^n
    \frac{ \rho_r^2 } { p_r }
  \right) \,
  \delta_{t, t'}
  .
\label{eq:zz_perfect}
\end{align}
Thus,
for $k \ge 2$, we get
\begin{equation}
  \kappa_k(t - t')
  =
  \sum_{i,j = 1}^n
  \phi_{ki} \phi_{kj}
  \langle \xi_i(t) \, \xi_j(t') \rangle
  =
  \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
  \, \delta_{t, t'},
\notag
\end{equation}
where we have used
Eqs. \eqref{eq:ortho1},
\eqref{eq:theta_def},
and
\eqref{eq:zz_perfect}.
Thus, Eq. \eqref{eq:kappa_delta} is exact,
with
$\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
In particular, if $p_i = \rho_i$, then, by
Eq. \eqref{eq:eig_orthonormal_rows}, we have
$\kappa_k(t - t') = \delta_{t, t'}$, or
\begin{equation}
  \Gamma_k = 1
  .
\label{eq:Gamma_perfect}
\end{equation}

\subsection{\label{sec:Gamma_onestep}
One-step process}

Another special case is the
one-dimensional one-step process\cite{vankampen},
which models a local sampling process.
In each step, the system hops to
one of the two neighboring bins
with equal probability $1/2$.
The transition matrix is
\begin{equation}
  \arraycolsep=3.6pt\def\arraystretch{1.4}
  \mathbf A
  =
  \left(
    \begin{array}{cccccccc}
      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
      \vdots & &  & & \vdots \\
      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
    \end{array}
  \right),
\notag
\end{equation}
where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
For many updating schemes
(cf. Sec. \ref{sec:band-matrix}),
we can further assume that
the matrices $\mathbf A$ and $\mathbf w$
share the eigenvectors,
with $\rho_i = p_i = 1/n$.
The eigenvalues of $\mathbf A$ are
$$
  \Lambda_k
  =
  \cos \frac{ (k - 1) \, g \, \pi }
            {            n        }
  ,
$$
where
$g = 2$ for a periodic variable, or $1$ otherwise.
In this case,
the autocorrelation function $\kappa_k(t)$
decays as $\kappa_k(0) \, \Lambda_k^t$,
with $\kappa_k(0) = 1$ being the same as
the value in perfect sampling.
%
Thus,
%
from Eq. \eqref{eq:Gamma_sum}, we get
\begin{equation}
  \Gamma_k
  =
  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
  =
  \frac{ 1 + \Lambda_k }
       { 1 - \Lambda_k }
  =
  \cot^2 \left[
    \frac{ (k - 1) \, g \, \pi }
         { 2 \, n }
  \right]
  .
\label{eq:Gamma_onestep}
\end{equation}

\subsection{\label{sec:Gamma_measure}
Measurement via an adaptive FDS simulation
}

For an unknown system,
we can use Eq. \eqref{eq:y2_eql} to
measure $\Gamma_k$'s in
an adaptive FDS simulation
with a constant schedule,
$\alpha(t) = a_0$.
However,
we often do not know a priori the intrinsic distribution, $p^*_i$,
and hence
cannot compute the deviation
of the bias potential, $V_i$.
To circumvent the problem, we
modify Eq. \eqref{eq:y2_eql} as
\begin{equation}
  \operatorname{var} Y_k
  =
  \frac{1}{2} \,
  a_0 \, \Gamma_k \, \lambda_k,
\label{eq:varY}
\end{equation}
where
``$\operatorname{var}$''
means the variance from the sampling (time) average,
and
\begin{align*}
Y_k
&=
\sum_{ i = 1 }^n
  \phi_{k i} \, (V_i - \bar V),
\end{align*}
with
$
\bar V
\equiv
\sum_{ j = 1 }^n \rho_j \, V_j.
$

To show Eq. \eqref{eq:varY},
we first note that
the difference, $V_i - {\bar V}$,
differs from the actual deviation, $x_i$,
defined in Eq. \eqref{eq:x_def},
only by a constant shift,
$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
Thus, the two share the same variance
under time averaging.
It follows that the corresponding linear combinations,
$Y_k$ and $y_k$, would also share the same variance.
But since the average of $y_k$ is zero,
we have
$$
\operatorname{var} Y_k
=
\operatorname{var} y_k
=
\left\langle y_k^2 \right\rangle.
$$
Finally, by using Eq. \eqref{eq:y2_eql},
we reach Eq. \eqref{eq:varY}.

As an example, we used the method to compute
the integrals, $\Gamma_k$'s, from
a local MC simulation
(cf. Sec. \ref{sec:results_system})
and an MD simulation,
and compared them with those from
the one-step sampling process.
As shown in Fig. \ref{fig:gamcmp},
the $\Gamma_k$'s of the small-$k$ eigenmodes
from the three sampling processes
were almost proportional,
suggesting that
the optimal schedule
for simulations using local sampling processes
would be similar in general properties.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
  }
  \caption{
    \label{fig:gamcmp}
    Integrals of the autocorrelation functions
    from a local MC simulation
    and an MD simulation.
    Each simulation was ran for $10^{10}$ steps.
    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
    For the MD simulation,
    the step size $\Delta t_\mathrm{MD} = 0.005$.
    The velocity rescaling thermostat\cite{bussi2007} was
    applied for the unit reduced temperature
    every MD step with viscosity $20$.
    The reflective boundary condition was used.
  }
\end{center}
\end{figure}

\section{\label{sec:procedure}
Procedure for computing the optimal schedule and error
}

Here we give a procedure for computing
the optimal schedule and the error
of a given adaptive FDS simulation.
Basically, we need to extract
$\lambda_i$'s and $\Gamma_i$'s
from the given system and updating scheme (input),
and then apply them to Eq. \eqref{eq:q_opt}
for the schedule,
as shown in Fig. \ref{fig:vardep}.
While the updating scheme alone
determines $\lambda_k$'s,
$\Gamma_k$'s depend also on the system
and the underlying sampling process (MC or MD).
By using the technique in Sec. \ref{sec:Gamma_measure}
to measure $\Gamma_k$'s,
we have the following procedure.

\begin{figure}[h]
\begin{singlespace}
  \fontsize{10pt}{\baselineskip}
  \begin{tikzpicture}
  \node[draw, text width={width("Updating ")}]
    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};

  \node[draw, text width={width("sampling")}]
    (sys) at (0, 1) {System\\and\\sampling\\process};

  \node[draw, text width={width("Eigenvalues,")}]
    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};

  \node[draw, text width={width("Integrals, ")}]
    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};

  \node[draw, text width={width("Simulation")}]
    (T) at (2.2, -0.5) {Simulation\\time, $T$};

  \node[draw, text width={width("Optimal")}]
    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};

  \node[draw, text width={width("Schedule,")}]
    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};

  \node[draw, text width={width("Error, $\Err$")}]
    (err) at (7, 0) {Error, $\Err$};

  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};

  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};

  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};

  \draw[->, >=latex, thick]
    (w) edge (lambda)
    (w) edge (gamma)
    (sys) edge (gamma);

  \draw[->, >=latex, thick]
    (lambda) -| (M1)
    (gamma)  -- (M1)
    (T)      -| (M1)
    (M1)     -- (R1)
    (R1)     |- (qT);

  \draw[->, >=latex, thick]
    (qT)     -| (R3);

  \draw[->, >=latex, thick]
    (R3) -- (N1) |- (alpha);

  \draw[->, >=latex, thick]
    (N1) |- (err);

  \draw[->, >=latex, thick]
    (R1) |- (R2) -| (R3);
\end{tikzpicture}
\end{singlespace}
\caption{
  \label{fig:vardep}
  Dependence of variables in computing
  the optimal schedule and the error.
}
\end{figure}

\begin{enumerate}

\item
From the updating matrix, $\mathbf w$,
compute the eigenvalues, $\lambda_k$'s.

\item
Run a preliminary adaptive FDS simulation
with a constant updating magnitude, $a_0$.

\item
Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.

\item
Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.

\item
Compute the optimal schedule by
numerically integrating Eq. \eqref{eq:q_opt}.

\item
Compute the error $E$ from
Eqs. \eqref{eq:error_split},
  \eqref{eq:error_res1},
  and
  \eqref{eq:error_asym2}.

\end{enumerate}

\section{\label{sec:stabilize_wband}
Stabilization of homogeneous updating schemes}

A homogeneous updating scheme
specified by an arbitrary kernel
is not necessarily stable,
i.e., some eigenvalues can be negative.
Here we show how to minimally modify
the updating kernel
to stabilize it.

Given an updating kernel,
we can compute the eigenvalues from
Eq. \eqref{eq:wband_eigenvalue}.
By replacing negative eigenvalues with zeros and
using the inversion formula,
Eq. \eqref{eq:mu_from_lambda_pbc}
or \eqref{eq:mu_from_lambda_refl},
we get a new updating kernel,
which is stable by construction.
However, the new kernel is widened
as $\mu_l$ may be nonzero for $l > b$.
If we truncate the new kernel at $\mu_b$,
the stability problem remains,
although the negative eigenvalues
tend to be smaller in magnitude.
Thus, to stabilize the updating kernel
without increasing the width,
we need to iterate the above process many times,
resulting the following algorithm.

\begin{enumerate}
  \item
    Given an updating kernel, $\mu_0, \dots, \mu_b$,
    compute the eigenvalues
    $\lambda_1, \dots, \lambda_n$
    from Eq. \eqref{eq:wband_eigenvalue}.
  \item
    If all eigenvalues are nonnegative within the numerical error,
    we are done. % and exit the loop.
  \item
    Otherwise, replace negative eigenvalues by zeros,
    and compute the new kernel,
    $\mu_0, \dots, \mu_b$, from
    Eq. \eqref{eq:mu_from_lambda_pbc}
    or
    \eqref{eq:mu_from_lambda_refl}
    with $\mu_l = 0$ for $l > b$.
    Go to step 1.
\end{enumerate}

\section{\label{sec:upperbound}
Asymptotic error for an updating scheme
with a set of well-separated eigenvalues}

We will derive here an upper bound of the asymptotic error
that can be associated to updating schemes
with a set of well-separated eigenvalues,
in parallel to the lower bound reached by
the single-bin scheme with a set of identical eigenvalues
(cf. Sec. \ref{sec:optWL}).

Starting from the definition Eq. \eqref{eq:mass_func}, we have
$$
  M(Q)
  \le
  \sum_{k = 2}^n
  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, Q }
  .
$$
Thus,
it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
\begin{equation}
  C
  =
  \int_0^\infty
  M(q') \, dq'
  \le
  \sum_{k = 2}^n \sqrt{ \Gamma_k }
  ,
  \label{eq:mint_ubound}
\end{equation}
and the asymptotic error from Eq. \eqref{eq:error_asym2}
is bounded as
\begin{equation}
  E_A
  =
  \frac{ C^2 } { T }
  \le
  \frac 1 T
  \left(
    \sum_{k=2}^n \sqrt{ \Gamma_k }
  \right)^2
  .
  \label{eq:error_asym_ubound}
\end{equation}
It is readily seen that this bound is always
greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
$$
\frac{T}{(T+t_0)^2} \, \sum_{k=2}^n \Gamma_k
\le
\frac{1}{T} \sum_{k=2}^n \Gamma_k
\le
\frac{1}{T} \left( \sum_{k=2}^n \sqrt{ \Gamma_k } \right)^2
.
$$

Next, we will show that this bound represents
updating schemes with a set of very different eigenvalues.
First, we have for any mode $k^*$ $(k^* \ge 2)$,
\begin{align}
  M(Q)
  &
  \ge
  \sqrt{ \Gamma_{ k^* } } \,
  \lambda_{ k^* } \,
  e^{ -\lambda_{ k^* } \, Q }
  ,
\label{eq:MQ_lbound1}
\end{align}
which would approach an equality if the $k^*$ mode
dominates the mass function, $M(Q)$,
defined in Eq. \eqref{eq:mass_func}.
%
Then, for the $k$th component of the asymptotic error,
we have from Eq. \eqref{eq:error_asym_Lagrangian}
\begin{align}
  \left\langle
  y_k^2(T)
  \right\rangle_A
  &=
  \int_0^T
  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
  \, {\dot q}^2(t) \, dt
  \notag
  \\
  &=
  \frac{ C } { T }
  \, \Gamma_k \, \lambda_k^2
  \int_0^{q(T)}
  \frac{ e^{-2 \, \lambda_k \, Q } } { M(Q) }
  \,  \, dQ
  \notag
  \\
  &\le
  \frac C T \,
  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
  \int_0^{q(T)}
    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, Q}
    \, dQ
  \notag
  \\
  &\le
  \frac C T \,
  \cdot
  \frac{ \Gamma_k }
  { \sqrt{ \Gamma_{k^*} } }
  \cdot
  \frac{ \lambda_k^2 }
  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
  ,
  \label{eq:y2_ubound_k}
\end{align}
where we have
used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
on the second and third lines, respectively,
and extended the upper limit to infinity on the last line.
For a set of well-separated eigenvalues,
$\lambda_k$ depends on the mode index $k$
much more critically than $\Gamma_k$.
Thus, the tightest upper bound
can be approximately obtained from minimizing the last factor
on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
and the minimal is reached at $k^* = k$,
when $\lambda_{k^*} = \lambda_k$.
%
So
%
\begin{equation}
  \left\langle
  y_k^2(T)
  \right\rangle_A
  \le
  \frac C T
  \sqrt{ \Gamma_k }
  .
  \label{eq:y2_ubound}
\end{equation}
By summing over $k$
and using Eq. \eqref{eq:mint_ubound},
we recover Eq. \eqref{eq:error_asym_ubound}.

This result shows that with a set of well-separated eigenvalues,
the asymptotic error profile would be
independent of $\lambda_k$'s and
be flatter than the initial one from
Eq. \eqref{eq:y2_eql}
(cf. Fig. \ref{fig:xerr}
for a Gaussian example).
Note, however, that to approach this upper bound
the simulation has to be long enough
such that the upper limit of the integral
can be extended from $q(T)$ to infinity
in the last step of Eq. \eqref{eq:y2_ubound_k}.
As a result,
the flatter profile from Eq. \eqref{eq:y2_ubound}
would stop at mode $k$,
at which $\lambda_k \approx 1/q(T)$.

\section{\label{sec:invt_schedule}
Modified inverse-time schedules}

Here we compute the error of the schedule given by
the modified inverse-time formula,
Eq. \eqref{eq:alpha_invtlambda}.

\subsection{\label{sec:invt_error}
General results on error
}

Using Eq. \eqref{eq:alpha_invtlambda}
in Eq. \eqref{eq:error_asym1} yields
the asymptotic error
\begin{align}
  \Err_A
  &=
  \frac{    1    }
       { T + t_0 }
  \sum_{k = 2}^n
    \frac{ \Gamma_k \, \nu_k^2 }
         {    2 \, \nu_k - 1   }
  \left[
    1 - \left(
          \frac {     t_0 }
                { T + t_0 }
        \right)^{ 2 \, \nu_k - 1 }
  \right]
  ,
\label{eq:error_asym_invt}
\end{align}
where $\nu_k \equiv \lambda_k / \lambda$.
From Eq. \eqref{eq:error_res1},
we get the residual error
\begin{equation}
  \Err_R
  =
  \frac { a_0 } { 2 }
  \sum_{k = 2}^n
  \Gamma_k \, \lambda_k \,
  \left(
      \frac{   t_0   }
           { T + t_0 }
  \right)^{ 2 \, \nu_k }
  .
\notag
\end{equation}
The total error is the sum of the two.

\subsection{\label{sec:invt_singlebin}
  Single-bin updating scheme
}

For the single-bin scheme, we have
$\lambda_2 = \cdots = \lambda_n = 1$,
and
$\Gamma_k$'s affect the error
only via the sum, $\sum_{k = 2}^n \Gamma_k$,
as a multiplicative constant.

In the asymptotic limit,
we can focus on the asymptotic error
and drop the second term in the square brackets
of Eq. \eqref{eq:error_asym_invt}
by assuming $\lambda < 2$.
%
Then,
%
$$
  E_A
  =
  \frac { 1 } { T + t_0 }
  \frac {             1            }
        { \lambda \, (2 - \lambda) }
  \sum_{ k = 2 }^n \Gamma_k
  ,
$$
is minimal at $\lambda = 1$.

\subsection{\label{sec:invt_nn}
Triple-bin updating scheme}

For the triple-bin updating scheme,
we can find an analytical expression
of the optimal value of $\lambda$
in the large-$n$ and asymptotic limit.
In this case,
if $2 \, \nu_k > 1$ for every $k$,
the asymptotic error is implicitly given by the integral
\begin{align}
  \Err_A
  =
  \frac{ 2 \, n  }
       { \pi \, T }
  \int_0^{\pi/2}
    \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
         {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
  \, dp
  .
\notag
\end{align}

For perfect sampling, if $\mathbf p = \pmb \rho$,
we have
$\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
and
$$
\begin{aligned}
  \Err_A
  &=
  \frac{n}{4 \, T}
  \left(
    \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
    +
    \frac{ \lambda }
    { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
  \right)
.
\end{aligned}
$$
Minimizing the function yields
\begin{equation}
  \lambda
  =
  \frac{ 1 - 4 \, \mu_1 }
       { 1 - 2 \, \mu_1 }
  ,
\label{eq:lambda_nn_perfect}
\end{equation}
which shows that the original inverse-time formula,
Eq. \eqref{eq:alpha_invt1},
is not optimal in this case.
The asymptotic error is given by
\begin{equation}
  \Err_A
  =
  \frac{n}{T}
  \left(
    1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
  \right)
  ,
\notag
\end{equation}
which
is minimal at $\mu_1 = 0$, the single-bin case.
Note that the value given by
Eq. \eqref{eq:lambda_nn_perfect}
was less than the observed optimal value
from Fig. \ref{fig:errinvt},
and the latter would decrease
with the simulation length.

For the one-step sampling process,
we find from Eq. \eqref{eq:Gamma_onestep}
that $\Gamma(p) = \cot^2 p$
peaks around $p \approx 0$,
%
Then
%
$$
  \Err_A
  \propto
  \frac{   2 \, n }
       { \pi \, T }
  \frac{             1             }
       {  \lambda \, (2 - \lambda) }
  .
$$
The optimal $\lambda$ is, therefore,
$\lambda \approx 1$.

\bibliography{simul}
\end{document}