### 2017-06-17 doc/response.tex
invt.tex: add back sec:seplambda

\documentclass{article}
\usepackage{amsmath}
\begin{document}

The old definition of $\rho_i$ and $p_i$ are assumed to be same now
to avoid unnecessary complexity, preserving the old $\rho_i$ only.
The old $p^*_i$ is replaced by the new $p_i$.
The old $V_i$ is replaced by $u_i$.
The old $x_i$ is replaced by $v_{*i}$.
The old $y_k$ is replaced by $\tilde v_k$.
The old $\theta_k$ is replaced by $\tilde f_k$.
The old $\zeta_i/p_i$ is replaced by $\zeta_i$.
The old $C$ is replaced by $C_M$.
The old degeneracy $g$ is replaced by $2/g$ for the multiple of period.
%
We have defined $f_i \equiv h_i/\rho_i$.
We changed $Q \to \bar q$ to imply the meaning of complement.

\end{document}