### 2017-06-28 doc/old/V7.1/invtV7.1.tex
adding Dr. Pettitt's V7.1

\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

\hypersetup{
  colorlinks,
  linkcolor={red!30!black},
  citecolor={green!20!black},
  urlcolor={blue!80!black}
}

\definecolor{DarkBlue}{RGB}{0,0,64}
\definecolor{DarkBrown}{RGB}{64,20,10}
\definecolor{DarkGreen}{RGB}{0,64,0}
\definecolor{DarkPurple}{RGB}{64,0,42}
\definecolor{LightGray}{gray}{0.85}
\newcommand{\repl}[2]{{\color{gray} [#1] }{\color{blue} #2}}
\newcommand{\add}[1]{{\color{blue} #1}}
\newcommand{\del}[1]{{\color{gray} [#1]}}
\newcommand{\note}[1]{{\color{DarkGreen}\footnotesize \textsc{Note.} #1}}
\newcommand{\answer}[1]{{\color{DarkBlue}\footnotesize \textsc{Answer.} #1}}
\newcommand{\summary}[1]{{\color{DarkPurple}\footnotesize \textsc{Summary.} #1}}

\newcommand{\Err}{E}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\bav}[1]{#1_\mathrm{av}}

\begin{document}

\title{Optimal updating magnitude in adaptive flat-distribution-sampling simulations}

\author{Cheng Zhang}
\author{Justin A. Drake}
\affiliation{
Sealy Center for Structural Biology and Molecular Biophysics,
The University of Texas Medical Branch,
Galveston, Texas 77555-0304, USA}
\author{Jianpeng Ma}
\affiliation{
Department of Biochemistry and Molecular Biology,
Baylor College of Medicine, Houston, Texas 77030, USA}
\affiliation{
Department of Bioengineering,
Rice University, Houston, Texas 77005, USA}
\author{B. Montgomery Pettitt}
\email{mpettitt@utmb.edu}
\affiliation{
Sealy Center for Structural Biology and Molecular Biophysics,
The University of Texas Medical Branch,
Galveston, Texas 77555-0304, USA}

\begin{abstract}
  We present a method of computing the optimal schedule
  of the updating magnitude
  for a class of flat-distribution sampling methods
  for free energy calculation including
  the Wang-Landau (WL) algorithm and metadynamics.
  These methods rely on adaptive construction of
  a bias potential that offsets
  the potential of mean force by histogram-based adaptive updates.
  The updating magnitude should decrease over time
  to reduce the error in the bias potential,
  and one may speed up the convergence by choosing an optimal schedule
  of the updating magnitude.
  We will show that
  the optimal schedule depends on the updating scheme.
  While the asymptotically optimal schedule for
  the single-bin updating scheme (commonly used in WL)
  is given by an inverse-time formula,
  that for the Gaussian updating scheme (commonly used in metadynamics)
  is often more complex.
  Further,
  we show that the single-bin updating scheme
  belongs to a class of bandpass updating schemes
  that are optimal for asymptotic convergence.
  These bandpass updating schemes aim at
  partially flattening a few selected histogram modes
  and their optimal schedule
  is also given by the inverse-time prescription.
  Constructed from orthogonal polynomials,
  bandpass updating schemes serve as generalizations
  of the Langfeld-Lucini-Rago algorithm
  and can be used for parameter tuning for the Gaussian ensemble method
  and umbrella sampling.
\end{abstract}

\maketitle

\section{Introduction}

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry
that can provide insight into an array of phenomena not easily studied
with traditional experiments.
Given a system,
often the task is to compute
a distribution, $p(z)$,
along a collective variable, $z$, by sampling via either Monte Carlo\cite{
  frenkel, newman, landau_binder} (MC)
or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
The dimensionless free energy, or potential of mean force (PMF),
along the collective variable, $z$,
is $-\ln p(z)$.
However, accurately estimating the PMF is often difficult
when the collective variable is energetically restricted to local regions
of the complicated, multimodal free energy surface.
Thus,
to overcome these energetic barriers and
to capture the global shape of the free energy surface
an effective strategy is to introduce a bias potential that
cancels the PMF.
Ideally, then, sampling yields a flat distribution
along $z$\cite{mezei1987, berg1992, *lee1993,
wang2001, wang2001pre,
huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}.
The flat-distribution sampling can be used to accelerate
sampling of slow transitions that otherwise would not readily
be observed in traditional simulations.

Many efficient flat-distribution-sampling (FDS) techniques
based on adaptive construction of the bias potential
have been introduced,
including the Wang-Landau (WL) algorithm\cite{
  wang2001, wang2001pre}
and metadynamics\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
among others\cite{kim2006, *kim2007, kim2010, junghans2014,
  langfeld2012, pellegrini2014,
  maragliano2006, *abrams2008,
  zheng2010}.
These techniques regularly update the bias potential
to discourage future visits to previously sampled configurations
by incrementally elevating the bias potential.
A key difference lies
in the updating window function,
which specifies
a neighborhood around the current value of
$z$ on the free energy surface
where the updating should occur
as well as the relative updating magnitude.
Often, the window function
takes the form of a discrete
$\delta$-function (a boxcar function one histogram bin wide)
in WL,
or that of a Gaussian
in metadynamics.\cite{junghans2014}
We will refer to the two updating schemes
as the single-bin
and Gaussian updating schemes, respectively.
While the former is popular for MC simulations\cite{wang2001,
  wang2001pre, kim2006, *kim2007},
the latter is more popular for MD simulations
as it provides a smoother bias potential
that can be readily differentiated for force calculation.\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
In both cases,
regular updates to the bias potential
disrupts the underlying equilibrium
sampling\cite{zhou2005, morozov2007, zhou2008},
and for convergence of the estimated PMF
one has to gradually decrease
the updating magnitude of the bias.

The schedule of reducing
the updating magnitude,
therefore, critically affects
the precision of the final bias potential,
and hence that of the PMF\cite{laio2005, bussi2006, poulain2006,
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007, min2007,
morozov2007, zhou2008,
komura2012, *caparica2012, *caparica2014,
barducci2008, dickson2011, dama2014}.
For the single-bin case, the optimal updating schedule
is a function of the
inverse time\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
morozov2007, zhou2008}
or
the inverse number of updating steps.

In this study,
we present a method of computing
the optimal schedule
for an adaptive FDS method
of a general updating scheme,
including the single-bin and Gaussian versions.
We will map the optimization problem to a mechanical one,
in which the schedule plays the role of the velocity of
a free particle with a position-dependent mass,
and the final error becomes the action.
In this way, the optimal schedule
can be deduced from the equation of motion
of the particle that minimizes the action.
The resulting optimal schedule
depends on the updating window function
through the mass function.
While the optimal schedule in the single-bin case
is given by the known inverse-time formula,
that for Gaussian case is more complex
and sensitive to the simulation length
and the width of the Gaussian window.
Further, we will show that
the single-bin updating scheme
is optimal for asymptotic convergence,
and it can be generalized
to a class of asymptotically optimal
bandpass updating schemes.
The latter allow one to partially flatten
a few long-range histogram modes,
and to avoid introducing local noise
into the bias potential during the updating process.
The optimal schedule of bandpass updating schemes
is also an inverse-time schedule.
Constructed from orthogonal polynomials,
bandpass updating schemes serve as
generalizations of the Langfeld-Lucini-Rago (LLR)
algorithm\cite{langfeld2012, pellegrini2014},
and can help other FDS
algorithms\cite{neuhaus2006, *neuhaus2007, zhu2012}
adjust the parameters of the bias potential automatically.

The article is organized as follows.
We present the analytical results in Sec. \ref{sec:theory},
provide numerical examples
in Sec. \ref{sec:results},
and conclude the article
in Sec. \ref{sec:conclusion}.

\section{\label{sec:theory}
Theory}

We develop the theory
in the following order.
In Sec. \ref{sec:background},
we briefly review the basics of FDS
and some known aspects of the optimal schedule.
Then, in Sec. \ref{sec:single-bin},
we derive the method of
computing the optimal schedule
in the simple case of the
single-bin updating scheme
used by the WL algorithm,
by proving the optimality
of the known inverse-time formula.
Next, we extend the method to the general case
of a multiple-bin updating scheme
that encompasses the Gaussian scheme used in metadynamics
in Sec. \ref{sec:multiple-bin}.
Finally, we compare different updating schemes
and show that the single-bin scheme
and a class of generalized bandpass schemes
are optimal for convergence
in the long time limit
in Sec. \ref{sec:cmpschemes}.

\subsection{\label{sec:background}
Flat-distribution sampling}

Consider the problem of computing
the distribution, $p_i$,
along a discrete quantity, $i$,
for a given system.
For a continuous quantity, $z$,
we can use the equivalent integer $i$ to represent
the index of a small interval, or a bin,
$(z, z + \delta z)$.

The difficulty of sampling a multimodal distribution
with deep valley(s) can sometimes be eased by
sampling a smoother biased distribution, $\rho_i$.
We will broadly define
a flat-distribution-sampling (FDS) simulation
as a biased simulation
targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$ for $n$ bins)
or a nearly flat one\cite{
dayal2004, *trebst2004, barducci2008, singh2011}.
The biased equilibrium distribution
under a bias potential, $u_i$, is given by
\begin{equation}
  \pi_i
  =
  \frac{                p_i \, e^{-u_i} }
       { \sum_{j = 1}^n p_j \, e^{-u_j} }
  \propto
  p_i \, e^{-u_i}
  ,
  \notag
\end{equation}
and our aim is to find a bias potential
such that the resulting $\pi_i$ would coincide
with the desired $\rho_i$.
Upon convergence, the bias potential would satisfy
\begin{equation}
  u_i \to \ln \frac{p_i}{\rho_i} + \mathrm{const.}
  ,
  \label{eq:Vi_target}
\end{equation}
and the PMF, $-\ln p_i$, can be readily
deduced from the bias potential. % $u_i$.

Many methods\cite{mezei1987, berg1992, *lee1993,
wang2001, wang2001pre, huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}
have been developed to find the proper bias potential.
In a class of equilibrium FDS methods\cite{
  mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
the bias potential, $u_i$, is
estimated a priori and fixed
during simulation,
and a correction to the bias potential
is derived from the normalized histogram, $H_i$,
accumulated from the simulation as
\begin{equation}
  \hat u_i
  =
  u_i
  +
  \ln \frac{ H_i }
           { \rho_i }.
  \label{eq:vcorr_equil}
\end{equation}
However, this formula often requires a long
accumulation period for a precise histogram,
and thus disallows
continuous improvement of the bias potential
as simulation lengthens.

Below we will focus on another class of adaptive FDS methods,
including the WL algorithm\cite{wang2001, wang2001pre}
and metadynamics\cite{huber1994, laio2002, *laio2008, *barducci2011, *sutto2012},
address the above shortcoming
by continuously and incrementally updating the bias potential.
In the WL algorithm\cite{wang2001, wang2001pre},
the bias potential, $u_i(t)$, is updated
at each step $t$ as
\begin{equation}
  u_i(t+1)
  =
  u_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { \rho_{i(t)} }
  ,
\label{eq:wl_update}
\end{equation}
where $i(t)$ is the bin at step $t$,
$\delta_{i, \, i(t)}$ is the Kronecker delta function,
which is $1$ when $i = i(t)$ or $0$ otherwise,
and $\alpha(t)$ is the updating magnitude.
We refer to this scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$.
In a more general \emph{multiple-bin scheme},
the updating also involves neighboring bins:
\begin{equation}
  u_i(t+1)
  =
  u_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) }
       { \rho_{i(t)} }
  .
  \label{eq:mbin_update}
\end{equation}
The Gaussian updating scheme used in metadynamics
is an example, with $w_{ij}$ being a Gaussian centered on $j$.
Note that if the updating occurs
only every few sampling steps,
then $t$ means the number of updating steps.

While adaptive FDS simulations
can often sample the desired distribution,
frequent updates to the bias potential
disrupt the underlying equilibrium sampling
and introduce artificial errors that need to be
reduced by gradually decreasing the updating magnitude\cite{
  belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
  zhou2005, morozov2007, zhou2008,
  laio2005, bussi2006, poulain2006, liang2007,
  crespo2010, *atchade2011, *fort2015}.

In the original WL algorithm\cite{
wang2001, wang2001pre},
the updating magnitude, $\alpha(t)$,
(or $\ln f$ therein)
is controlled in stages.
In each stage, $\alpha(t)$
is kept constant,
and the histogram along $i$
is collected and monitored.
Once the histogram is sufficiently flat,
we switch to a new stage
with $\alpha(t)$ reduced by a factor of $1/2$\cite{
wang2001, wang2001pre}.
While this scheme works well for early stages,
it tends to reduce the magnitude
too quickly in late stages, making the asymptotic error
saturate\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.

A more effective way
of controlling the schedule, $\alpha(t)$,
in late stages
is to follow the formula
\begin{equation}
  \alpha(t) = \frac{1}{t},
  \label{eq:alpha_invt}
\end{equation}
where $t$,
referred to as the ``time'' below,
is the number of updating steps
from the beginning of the simulation\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
morozov2007, zhou2008,
komura2012, *caparica2012, *caparica2014}.
This general result\cite{
  robbins1951, pellegrini2014}
has an intuitive explanation\cite{
  marsili2006, barducci2008}
that interprets the bias potential
as a runtime average computed from all data points collected so far,
which absorbs the contribution of the newest data point
by its due weight,
the inverse of the current sample size, $t$
(cf. Appendix \ref{sec:equilerr}).

\subsection{\label{sec:single-bin}
Single-bin updating scheme}

In this section,
we will derive the optimal $\alpha(t)$
for the single-bin scheme,
Eq. \eqref{eq:wl_update},
as a preparation
for the more general multiple-bin scheme.
We will first express the error of
the bias potential, $u_i(t)$,
as a functional of $\alpha(t)$,
and then minimize it by variation.

\subsubsection{Error}

We define the net error in the bias potential
by deducting two contributions.
First, to deduct the target value given by
Eq. \eqref{eq:Vi_target},
we introduce a shifted bias potential
\begin{equation}
  v_i(t)
  \equiv
  u_i(t)
  -
  \ln \frac { p_i }
            { \rho_i }
  ,
  \notag
\end{equation}
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i(t)$ are
equivalent to the updates to $u_i(t)$,
since the shift
remains a constant during simulation.

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
We define the bias potential deviation from the
bin average, $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
as
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  .
\label{eq:x_def}
\end{equation}
Below we will use the notations, $x_{*i}$ and $\bav{x}$,
for a general variable, $x_i$.
The total error is then given by
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \,
  \left\langle v_{*i}^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

\subsubsection{\label{sec:sbin_diffeq}
Differential equation}

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation\footnote{In
going to the continuous-time setup,
we find it more convenient to shift the origin of time by $-1$,
e.g. the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
\begin{equation}
  \dot u_i(t)
  \equiv
  \frac{ d u_i(t) } { dt }
  \approx
  \alpha(t) \, \frac{ h_i(t) } { \rho_i }
  \equiv
  \alpha(t) \, f_i(t)
  ,
  \label{eq:ut_diffeq}
\end{equation}
with
$h_i(t) = \delta_{i, i(t)}$
being the instantaneous histogram,
which is equal to $1$
for the bin $i(t)$ at time $t$
or zero otherwise,
and we have defined
$f_i(t) \equiv h_i(t) /\rho_i$.
Clearly,
\begin{equation}
  \bav{f}(t) = \sum_{i=1}^n h_i(t) = 1
  ,
  \label{eq:fav1}
\end{equation}
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.
Deducting the bin average [cf. Eq. \eqref{eq:x_def}],
we get
\begin{equation}
  \dot v_{*i}(t)
  \approx
  \alpha(t) \, f_{*i}(t)
  .
  \label{eq:vt_diffeq}
\end{equation}

We can split the histogram into
a deterministic averaged part, $\langle f_i(t) \rangle$,
and a random fluctuating part, $\zeta_i(t)$,
\begin{equation}
  f_i(t) =
  \langle f_i(t) \rangle
  +
  \zeta_i(t)
  .
  \notag
\end{equation}
The deterministic part can be related
to an average in an ensemble consisting of
many copies of similar simulations
sharing the same schedule $\alpha(t)$
and the same bias potential at time $t$.
The initial states and the stochastic forces
during the process may differ, however.
For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period,
and we may assume a quasi-equilibrium sampling process
such that
\begin{align}
  \langle f_i(t) \rangle
  &\approx
  \frac{ \pi_i(t) } { \rho_i }
  =
  \frac{                          e^{-v_{*i}(t)} }
       { \sum_{j = 1}^n \rho_j \, e^{-v_{*j}(t)} }
  \approx
  1 - v_{*i}(t)
  ,
  \notag
\end{align}
where we have assumed the smallness
of $v_{*i}(t)$ in the linear expansion.
Deducting the bin average and using Eq. \eqref{eq:fav1}, we get
$\langle f_{*i}(t) \rangle \equiv \langle f_i(t) \rangle - 1 \approx - v_{*i}(t)$,
and
\begin{align}
  f_{*i}(t)
  &\approx
  - v_{*i}(t)
  +
  \zeta_i(t)
  .
  \label{eq:sh_ave}
\end{align}

We will approximate the histogram fluctuation part, $\zeta_i(t)$,
hence its deviation from the bin average, $\zeta_{*i}(t)$,
as white noise such that
\begin{equation}
  \langle \zeta_{*i}(t) \, \zeta_{*i}(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \notag
\end{equation}
for some $G_i$,
and $\delta(t)$ is the Dirac delta function.
We expect the approximation to hold for simulation
much longer than the autocorrelation time.

From Eqs.
\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
we get % a set of decoupled equations
\begin{equation}
  \dot v_{*i}(t)
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \zeta_{*i}(t) \right]
  ,
  \notag
\end{equation}
and the formal solution is
\begin{equation}
  v_{*i}(T)
=
  v_{*i}(0) \, e^{-q(T)}
+
\int_0^T
  \dot \vartheta\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
\notag
\end{equation}
where
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
and
\begin{align}
\vartheta(q')
&\equiv
e^{q' - q(T)}.
\label{eq:theta_def}
\end{align}

\subsubsection{Optimal schedule}

At the end of period $T$,
the total error defined in Eq. \eqref{eq:error_def}
is a sum of a residual error, $E_R$ (from the decay of the initial error)
and an asymptotic error, $E_A$:
\begin{align}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T)
  ,
  \label{eq:error_tot}
\end{align}
where
\begin{align}
  \Err_R(T)
  &= \Err(0) \, e^{-2 \, q(T)}
  ,
  \label{eq:ER_sbin}
  \\
  \Err_A(T)
  &= \Gamma \, \int_0^T \dot \vartheta^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:EA_sbin}
\end{align}
with
$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$,
via $v_{*i}(T)$.
If $\alpha(t)$ is a constant, $a_0$,
so that $q(t) = a_0 \, t$,
then after a long period,
the error approach an equilibrium value
\begin{equation}
  \Err_\mathrm{equil.}
  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
  = \frac{1}{2} a_0 \Gamma
  .
  \label{eq:error_equil_sbin}
\end{equation}
Our aim is to find the $q(t)$
that minimizes this error.
We now fix the endpoint value, $q(T)$,
and vary the curve, $q(t)$ ($0 < t < T$).
This is equivalent to varying the curve,
$\vartheta\bigl( q(t) \bigr)$,
with fixed endpoints,
$\vartheta\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$\vartheta\bigl( q(T) \bigr) = 1$.
Since the residual error
is fixed during the process,
the asymptotic error,
resembling the action of a free particle,
demands a constant velocity
\begin{equation}
  \dot \vartheta\bigl( q(t) \bigr) = c
  .
\label{eq:dthetadt_const}
\end{equation}
Then, by using Eq. \eqref{eq:theta_def}, we have
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
where $t_0$ is some constant,
and $c$ has been determined
from the $t = T$ case
in the second step.
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
and Eq. \eqref{eq:alpha_invt}
is the special case of $t_0 = 0$.

We now determine the optimal value of $q(T)$,
which is fixed in the above.
This is equivalent to choosing a value of $t_0$
\big[because $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$,
from the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
Using Eq. \eqref{eq:alpha_invt1} in
Eqs. \eqref{eq:ER_sbin} and \eqref{eq:EA_sbin}
yields
\begin{align}
  \Err_R(T)
  &= \frac{ \Err(0) \, t_0^2 } { (T + t_0)^2 }
  ,
  \notag
  \\
  \Err_A(T)
  &= \frac{ \Gamma \, T } { (T + t_0)^2 }
  ,
  \label{eq:EA_sbin1}
\end{align}
and the total has a minimum of
\begin{equation}
  \Err(T)
  =
  \frac{ \Gamma } { T + t_0 }
  ,
  \label{eq:Emin_sbin}
\end{equation}
which is reached at $t_0 = \Gamma /\Err(0)$,
or if we assume Eq. \eqref{eq:error_equil_sbin} for $\Err(0)$,
\begin{equation}
  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
  .
  \label{eq:t0_sbin}
\end{equation}
The minimum error can be related to
the expected histogram fluctuation\cite{zhou2005, zhou2008}
(cf. Appendix \ref{sec:hfluc}).

\subsection{\label{sec:multiple-bin}
Multiple-bin updating scheme}

We now turn to a general multiple-bin updating scheme.
By definition, Eq. \eqref{eq:mbin_update},
a visit to bin $j$ in this case results in updates of the bias potential
not only at bin $j$, but also at a few neighboring bins $i$'s.
Similar to Eq. \eqref{eq:ut_diffeq},
we can approximate the updating scheme
as a differential equation,
\begin{equation}
  \dot u_i(t)
  \approx
  \alpha(t) \,
  \sum_{j=1}^n w_{ij} \, f_j(t)
  =
  \alpha(t) \,
  \sum_{j=1}^n w_{ij} \, \frac{ h_j(t) } { \rho_j }
  .
  \label{eq:ut_diffeq_mbin}
\end{equation}
The optimization procedure is similar to
the single-bin case.
The key simplification comes from the
projection of the bias potential to the $n$ eigenmodes
of the updating matrix, $\mathbf w$,
formed by the relative magnitudes, $w_{ij}$'s.

\subsubsection{\label{sec:updating-matrix}
Updating matrix and eigenmode decomposition}

The matrix, $\mathbf w$, is subject to several conditions.
First, we have a fixed-point condition\cite{bussi2006, dama2014}.
To sample the desired distribution,
$\pmb\rho = (\rho_1, \dots, \rho_n)$,
the growth rate of the bias potential,
${\dot u}_i(t)$
in Eq. \eqref{eq:ut_diffeq_mbin},
should be a constant of $i$
if $h_j(t)$ were the same as $\rho_j$,
i.e.
$\sum_{j=1}^n w_{ij}$ should be a constant
to allow the bias potential to grow uniformly
in the asymptotic regime.
By a proper overall scaling of $\mathbf w$,
we may write this condition as
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1
  .
\label{eq:w_sumj}
\end{equation}
In other words, $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
except that some elements can be negative
in our cases.
Below we will %take advantage of the analogy and
borrow techniques used
in studying transition matrices\cite{vankampen},
but limit ourselves to properties that are unaffected
by the admission of negative elements into $\mathbf w$.

Second, we limit ourselves to matrices % $\mathbf w$'s
that satisfy
the detailed balance condition for $\pmb\rho$,
\begin{equation}
  \rho_i \, w_{ij} = \rho_j \, w_{ji}
  .
  \label{eq:w_detailedbalance}
\end{equation}
This requires the scaled updating matrix,
$\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
to be symmetric,
and the symmetry allows the diagonalization
of $\mathbf w$ with a set of
eigenvectors, $\phi_{ki}$,
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij}
  =
  \lambda_k \, \phi_{kj}
  ,
\label{eq:eig_w}
\end{equation}
satisfying the orthonormal conditions\cite{vankampen}:
\begin{align}
  \sum_{k = 0}^{n - 1}
    \phi_{ki} \, \phi_{kj}
  &=
  \delta_{ij} \, \rho_i,
  \label{eq:eig_orthonormal_cols}
  \\
  \sum_{i = 1}^n
    \frac{ \phi_{ki} \, \phi_{li} }
         { \rho_i }
  &=
  \delta_{kl}
  ,
  \label{eq:eig_orthonormal_rows}
\end{align}
where we have set the index of the first eigenmode to $0$
instead of $1$ for later convenience.

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \notag
\end{equation}
We will label this eigenvector by $k = 0$,
and
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
with $\lambda_0 = 1$.
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho0}
\end{equation}
We may reconstruct the updating matrix in $\phi_{ki}$'s
as\cite{bussi2006}
\begin{equation}
  w_{ij}
  =
  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
  \lambda_k \, \phi_{ki} \, \phi_{kj}
  .
  \label{eq:w_from_phi}
\end{equation}

We can further define
a generalized Fourier transform, $\mathcal{F}$,
for variable, $x_i$,
\begin{equation}
  {\tilde x}_k
  \equiv \mathcal{F}[x_i]_k
  \equiv \sum_{i = 1}^n \phi_{ki} \, x_i
  ,
  \notag
\end{equation}
and the inverse transform is
$x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
Then,
we can rewrite Eq. \eqref{eq:ortho0} as
$\mathcal F[1]_k = \delta_{k0}$,
and Eq. \eqref{eq:eigenmode0} means that
the bin average
$\bav{x}= \mathcal F[x_i]_0 = \tilde x_0$,
and
\begin{equation}
  \tilde x_{*k}
  = \tilde x_k - \bav{x} \, \delta_{k0}
  =
  \begin{dcases}
    0           & k = 0
    ,
    \\
    \tilde x_k  & k > 0
    .
  \end{dcases}
  \label{eq:xtstar}
\end{equation}
Thus, the only effect
of the bin-average deduction on $\tilde x_k$ is to
annihilate the $k=0$ mode.
We also have Parseval's theorem,
\begin{equation}
  \sum_{k=1}^{n-1} \tilde x_k \, \tilde y_k
  =
  \sum_{k=0}^{n-1} \tilde x_{*k} \, \tilde y_{*k}
  =
  \sum_{i=1}^n x_{*i} \, y_{*i}
  \label{eq:parseval}
\end{equation}

We can now use Eq. \eqref{eq:eig_w} to diagonalize
the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
as
\begin{equation}
  {\tilde v}_k(t + 1) =
  {\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
  {\tilde f}_k(t)
  ,
  \label{eq:vkupdate}
\end{equation}
or in continuous time, we have, for $k > 0$,
\begin{equation}
  \dot{\tilde v}_k(t)
  =
  \alpha(t) \, \lambda_k \, {\tilde f}_k(t)
  \approx
  -\alpha(t) \, \lambda_k \,
  \bigl[ {\tilde v}_k(t) - {\tilde \zeta}_k(t) \bigr]
  ,
  \label{eq:vt_diffeq_mbin}
\end{equation}
where
we have used Eq. \eqref{eq:xtstar} and
the transformed version of Eq. \eqref{eq:sh_ave}
in the second step.

Equation \eqref{eq:vt_diffeq_mbin}
shows that the rate of error reduction
is given by $\alpha(t) \, \lambda_k$.
Thus, a \emph{stable} updating scheme
permits no negative eigenvalue, $\lambda_k$,
for $k > 0$.
With a set of heterogeneous eigenvalues,
the errors of the eigenmodes are reduced at different rates.
Thus, the optimization problem of the final error
becomes more complicated and it often
involves a compromise among eigenmodes.

\subsubsection{Error}

By Eq. \eqref{eq:parseval},
we can rewrite the error defined in Eq. \eqref{eq:error_def} as
$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
As in Eq. \eqref{eq:error_tot},
we can decompose the error as the sum
of the residual error and the asymptotic error, with
\begin{align}
  \Err_R(T)
  &=
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) }
  ,
  \label{eq:error_res}
  \\
  \Err_A(T)
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot \vartheta_k^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:error_asym}
\end{align}
where
\begin{equation}
  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  ,
  \label{eq:uk_def}
\end{equation}
and $\Gamma_k$ is computed from
the autocorrelation function
of the histogram fluctuation,
$\bigl\langle {\tilde \zeta}_k(0)
\, {\tilde \zeta}_k(t) \bigr\rangle$,
as:
\begin{equation}
  \Gamma_k
  = \sum_{t = -\infty}^\infty
  \bigl\langle {\tilde \zeta}_k(0)
            \, {\tilde \zeta}_k(t) \bigr\rangle
  .
  \label{eq:Gamma_sum}
\end{equation}
Equation \eqref{eq:Gamma_sum} represents
a first approximation for modeling
a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
In Appendix \ref{sec:Gamma_measure},
we discuss a way of estimating $\Gamma_k$
from an FDS simulation under a constant magnitude.

To model the initial error for Eq. \eqref{eq:error_res},
we will assume that
the preliminary adaptive FDS sampling has been
performed under a constant updating magnitude for $T_0$ steps.
If the initial bias potential
before the preliminary run is zero,
one can show that
\begin{equation}
  \left\langle
    {\tilde v}_k^2(0)
  \right\rangle
  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
  + \epsilon_k
  ,
  \label{eq:xt2_eql1}
\end{equation}
where
$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$
and if no eigenvalue is zero,
then we can ignore the $\epsilon_k$ term
in the case of $T_0 \to \infty$,
\begin{equation}
  \left\langle
    {\tilde v}_k^2(0)
  \right\rangle
  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
  .
  \label{eq:xt2_eql}
\end{equation}

\subsubsection{\label{sec:optschedule}
Optimal schedule}

As in the single-bin case,
by varying the schedule, $\alpha(t)$, under a fixed value of
$q(T) = \int_0^T \alpha(t) \, dt$,
we can focus on the asymptotic error,
$\Err_A(T)$.
The value of $q(T)$ will be adjusted later
to minimize the total error in Sec. \ref{sec:optinitalpha}).
We can rewrite Eq. \eqref{eq:error_asym} much like an action
of a particle whose position is given by $q(t)$:
\begin{equation}
  \Err_A(T)
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
  \notag
\end{equation}
where the Lagrangian is
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 1 }^{n - 1}
    \Gamma_k \, {\dot \vartheta}_k^2\bigl[ q(t) \bigr]
  =
  M^2\bigl(q(T) - q(t) \bigr)
  \; \dot q^2( t )
  .
\notag
\end{align}
where we have
used Eq. \eqref{eq:uk_def} and
defined the (square-root) mass function as
\begin{equation}
  M(\bar q)
  \equiv
  \sqrt{
    \textstyle\sum_{ k = 1 }^{n - 1}
    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, \bar q}
  }
  .
  \label{eq:mass_func}
\end{equation}
In this mechanical analogy,
the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
and the above Lagrangian, containing only the kinetic energy,
describes a free particle
with a position-dependent mass.
Here, the Hamiltonian coincides with the Lagrangian:
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L
  .
  \notag
\end{equation}
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold}
which means that the asymptotic error grows
linearly with time at a rate of $\mathcal L$.
We may set
\begin{equation}
  \sqrt{ \mathcal H }
  =
  \sqrt{ \mathcal L }
  =
  M\bigl( q(T) - q(t) \bigr)
  \;
  \dot q(t)
  =
  \frac{C_M}{T}
  ,
  \label{eq:Lagrangian_const}
\end{equation}
for some positive $C_M$
as a generalization of Eq. \eqref{eq:dthetadt_const}.
By integrating this equation of motion, we get
\begin{equation}
  \int_{ q(T) - q(t) }^{ q(T) }
    M(\bar q)
    \;
    d \bar q
  =
  C_M \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
and $C_M$ can be determined from
the $t = T$ case as
\begin{equation}
  C_M =
  \int_{ 0 }^{ q(T) }
    M( \bar q )
    \;
    d \bar q
  .
  \label{eq:mint}
\end{equation}
Thus, Eq. \eqref{eq:Lagrangian_const}
gives an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
The minimal asymptotic error is
\begin{equation}
  \Err_A(T)
  =
  \mathcal L \, T
  =
  \frac { C_M^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(\bar q) \, d \bar q
  \right)^2
  .
\label{eq:error_asym2}
\end{equation}

\subsubsection{\label{sec:mass_distr}
Characterization of optimal schedules}

We can characterize the optimal schedule
geometrically as follows.
First, we define
a normalized mass distribution as
\begin{equation}
  m(\bar q)
  =
  \frac{
    M(\bar q)
  }
  {
    \int_0^{ q(T) } M(q') \, d q'
  }
  =
  \frac{
    M(\bar q)
  }
  {
    C_M
  }
  ,
\notag
\end{equation}
such that
$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
we find that for $\bar q(t) \equiv q(T) - q(t)$,
\begin{align}
  m\left( \bar q(t) \right)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
  \label{eq:mQ_invTa}
  \\
  \int_{\bar q(t)}^{ q(T) }
    m(q') \, d q'
  &=
  \frac t T
  ,
  \label{eq:intmQ_tT}
\end{align}
Thus, each point on the schedule,
$\bigl(t, \alpha(t)\bigr)$,
can be mapped to a point,
$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
on the mass distribution,
such that the ordinate of the latter curve
is $1/(T\alpha)$,
and the area under the curve in the domain, $[\bar q(t), q(T)]$,
is equal to $t/T$,
as shown in Fig. \ref{fig:massq}(a).

\begin{figure}[h]
\begin{center}
  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
    \node [] at (-0.3, 2.0) {(a)};

    \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
      plot({\x}, {2*exp(-\x)});

    \fill [gray!30!white, domain=0.7:1.609, variable=\x]
      (0.7, 0) --
      plot ({\x}, {2*exp(-\x)})
      -- (1.609, 0) -- cycle;

    \draw[domain=0:1.609, variable=\x, very thick]
      plot({\x}, {2*exp(-\x)});

    \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};

    \draw[<->] (0.6, 0)  --
               node[left, inner sep=1mm]
               {$\dfrac{1}{T \, \alpha}$}
               (0.6, 0.99317);

    \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
    \draw[->] (0, 0) -- (2.2, 0);
    \node[] at (2.2, -0.2) {$\bar q$};
    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);

    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
    \draw[] (1.609, 0.1) -- (1.609, 0.0) node[below] {$q(T)$};
  \end{tikzpicture}
  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
    \node [] at (-0.3, 2.0) {(b)};
    \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
      plot({\x}, {1/(\x+0.5)});

    \fill [gray!30!white, domain=0.7:2.0, variable=\x]
      (0.7, 0) --
      plot ({\x}, {1/(\x+0.5)})
      -- (2.0, 0) -- cycle;

    \draw[domain=0:2.0, variable=\x, very thick]
      plot({\x}, {1/(\x+0.5)});

    \node[] at (1.1, 0.35) {$\bar q(t)$};

    \draw[<->] (0.6, 0)  --
               node[left, inner sep=1mm]
               {$\dfrac{1}{T \, m}$}
               (0.6, 0.83333);

    \draw[] (0.5, 0.83333) -- (0.7, 0.83333);
    \draw[->] (0, 0) -- (2.6, 0);
    \node[] at (2.6, -0.2) {$t$};
    \draw[->] (0, 0) -- node[above, rotate=90] {$\alpha(t)$} (0, 2.2);

    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$t$};
    \draw[] (2.0, 0.1) -- (2.0, 0.0) node[below] {$T$};
  \end{tikzpicture}
  \caption{
    \label{fig:massq}
    (a) Geometric construction of the optimal schedule,
    $\alpha(t)$,
    from the normalized mass distribution,
    $m(\bar q)$.
    (b) Inverse construction of $m(\bar q)$
    from the optimal schedule $\alpha(t)$.
  }
\end{center}
\end{figure}

The schedule, $\alpha(t)$,
also defines a variable transformation from
the mass distribution, $m(\bar q)$.
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited to the interval $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$,
as shown in Fig. \ref{fig:massq}.
A few examples are shown in Table \ref{tab:m_and_a}.
For example, in the single-bin scheme,
the mass distribution decays exponentially,
and we would have the inverse-time schedule.
On the other hand,
if the mass distribution is inversely proportional to $\bar q$,
which is approximately true for the Gaussian updating scheme
at large $\bar q$,\footnote{If
  the magnitudes of the eigenvalues, $\lambda_k$'s,
  span a wide range, then for a large $\bar q$,
  the sum in Eq. \eqref{eq:mass_func} will be dominated by
  the largest term, where $\lambda_k \approx 1/\bar q$,
  and $M(\bar q) \approx \sqrt{\Gamma_k}/\bar q$.
  }
then we would have an exponentially decaying schedule.

\begin{table}[h]
  \caption{\label{tab:m_and_a}
    Examples of the variable transformation between the
    reduced schedule, $\alpha^*(t^*)$,
    and the reduced mass distribution, $m^*(\bar q^*)$,
    as defined in Eqs. \eqref{eq:mQ_invTa}
    and \eqref{eq:intmQ_tT}.
  }
  \setlength{\tabcolsep}{2pt} % spacing between columns
  \renewcommand\arraystretch{2.0} % spacing between rows
  \begin{tabular} { c c c }
    \hline
    $m^*(\bar q^*)$ &
    $\alpha^*(t^*)$ &
    $\bar q^*(t^*)$
    \\
    \hline
    $\dfrac{1}{\gamma \, ({\bar q}^* + s_\gamma) }$ &
    $(1 + s_\gamma) \, \gamma \, e^{-\gamma \, t^*}$ &
    $(1 + s_\gamma) \, e^{-\gamma \, t^*} - s_\gamma$
    \\
    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
    \\
    $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
    $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{1/a} - 1\right]$
    \\
    \hline
    \multicolumn{3}{p{8cm}}{
    Here,
    $\gamma > 0$,
    and we have defined
    $s_\gamma \equiv 1/(e^\gamma - 1)$,
    ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
    and
    $c \equiv {q^*_0}^a(1 + t^*_0)$.
    } \\
    \hline
  \end{tabular}
\end{table}

\subsubsection{\label{sec:optinitalpha}
  Optimal initial updating magnitude
}

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
We will consider simulations in which
the system is initially equilibrated
at a constant updating magnitude, $a_0$.
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
\begin{align}
  \mathcal Y\bigl( q(T) \bigr)
  &\equiv
  \frac{ M\bigl( q(T) \bigr) } { T }
    \int_0^{q(T)} M(q') \, d q'
  -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  \notag \\
  &
  -\sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  =0
  .
\label{eq:opt_qT}
\end{align}
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, wang2001pre}.
}

\subsubsection{\label{sec:band-matrix}
Homogeneous updating schemes}

Many commonly-used updating schemes,
such as the Gaussian updating scheme
used in metadynamics,
can be characterized by a rigid window
centered on the current bin.
The eigenvalues of these translationally invariant or homogeneous
schemes can be readily deduced
from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
as a cosine transform,
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ i = 1 }^b
  \mu_i
  \cos\left(
  \frac{ k \, i \, 2 \, \pi }
       {      g \, n        }
  \right)
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
where $g = 1$ here for the periodic variable,
or $2$ for a non-periodic one
(cf. Appendix \ref{sec:more_wband}).
For the Gaussian updating scheme,
we have\cite{bussi2006}
\begin{align}
  \lambda_k
  =
  \exp\left[
        -\frac 1 2
        \left(
         \frac{ 2 \, \pi \, k \, \sigma_z }
              { g \, \Delta z  }
        \right)^2
      \right]
  ,
\label{eq:lambda_Gaussian}
\end{align}
where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
are the width of the Gaussian and the span
in the original collective variable, $z$,
respectively.

\subsubsection{\label{sec:procedure}
Procedure of computing the optimal schedule and error
}

Based on the above results,
we now give a procedure of computing
the optimal schedule and the error
for a given adaptive FDS simulation.
Basically, we need to extract
$\lambda_k$'s, $\Gamma_k$'s, and $\epsilon_k$'s
from the given system and updating scheme (input),
and then apply them to Eq. \eqref{eq:q_opt}
for the schedule,
as shown in Fig. \ref{fig:vardep}.
While the updating scheme alone
determines $\lambda_k$'s,
$\Gamma_k$'s depend also on the system
and the underlying sampling process.
We have

\begin{figure}[h]
\begin{singlespace}
  \fontsize{10pt}{\baselineskip}
  \begin{tikzpicture}
  \node[draw, text width={width("Updating ")}]
    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};

  \node[draw, text width={width("sampling")}]
    (sys) at (0, 1.1) {System\\and\\sampling\\process};

  \node[draw, text width={width("Eigenvalues,")}]
    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_1, \dots, \lambda_{n-1}$};

  \node[draw, text width={width("Integrals, iii")}]
    (Gamma) at (2.1, 1.1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$\\and biases,\\$\epsilon_1, \dots, \epsilon_{n-1}$};

  \node[draw, text width={width("Simulation")}]
    (T) at (2.2, -0.5) {Simulation\\time, $T$};

  \node[draw, text width={width("Optimal")}]
    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};

  \node[draw, text width={width("Schedule,")}]
    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};

    \node[draw, text width={width("Error,")}]
    (err) at (7, 0) {Error, $\Err(T)$};

  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};

  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};

  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};

  \draw[->, >=latex, thick]
    (w) edge (lambda)
    (w) edge (Gamma)
    (sys) edge (Gamma);

  \draw[->, >=latex, thick]
    (lambda) -| (M1)
    (Gamma)  -- (M1)
    (T)      -| (M1)
    (M1)     -- (R1)
    (R1)     |- (qT);

  \draw[->, >=latex, thick]
    (qT)     -| (R3);

  \draw[->, >=latex, thick]
    (R3) -- (N1) |- (alpha);

  \draw[->, >=latex, thick]
    (N1) |- (err);

  \draw[->, >=latex, thick]
    (R1) |- (R2) -| (R3);
\end{tikzpicture}
\end{singlespace}
\caption{
  \label{fig:vardep}
  Computation of the optimal schedule and the error.
}
\end{figure}

\begin{enumerate}

\item
Compute the eigenvalues, $\lambda_k$'s
from the updating matrix, $\mathbf w$,
or from Eq. \eqref{eq:wband_eigenvalue}
for a homogeneous updating scheme.
Note that
no eigenvalue can be negative for the updating scheme to be stable.

\item \label{step:prerun}
Run a preliminary adaptive FDS simulation
under a constant updating magnitude, $a_0$.

\item \label{step:Gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
(cf. Appendix \ref{sec:Gamma_measure})
as well as the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it leads to overestimated
  $\epsilon_k$ for large-$k$ modes.
}

\item \label{step:qT}
Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
We implemented a variant of the Newton-Raphson method\cite{press3rd}
with the bisection method\cite{press3rd}
as a fallback.\footnote{
The equation was solved for $\ln q(T)$
to avoid negative values of $q(T)$,
and we imposed a lower bound,
$\bigl|\mathcal Y\bigl(q(T)\bigr)/q(T)\bigr|$, for
$\mathcal Y'\bigl(q(T)\bigr)$
in computing its change.
The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
}

\item \label{step:alpha}
Compute the optimal schedule by
  numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
To handle Gaussian updating schemes more effectively,
  we have chosen the $\mathcal N+1$ grid points
  ($\mathcal N=20000$ here)
  over $[0, q(T)]$ to be given by
  $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
  where $i = 0, \dots, \mathcal N$.\footnote{For
    a mass distribution, $m(\bar q)$, that is
    inversely proportional to $\bar q$,
    this grid would make the integral,
    or the difference, $t_{[i+1]} - t_{[i]}$,
    roughly a constant.}
The curve $q(t)$ is then integrated
  using the trapezoidal rule\cite{press3rd},
  resulting in a list of values, $\bigl(q_{[i]}, t_{[i]} \bigr)$.
We then compute the updating magnitude by numerical differentiation:
  $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$\footnote{This
    discretization follows from runtime average interpretation
    discussed in Appendix \ref{sec:equilerr}.}
  and
  $\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
  with
  $q_{[ -1]} \equiv 0$,
  $q_{[N+1]} \equiv q(T)$,
  $t_{[ -1]} \equiv 0$, and
  $t_{[N+1]} \equiv T$.
The resulting values of $(t_{[i]}, \alpha_{[i]})$
  can be linearly interpolated
  to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.

\item
  Compute the total error $\Err(T)$ from
Eqs. \eqref{eq:error_tot},
  \eqref{eq:error_res},
  \eqref{eq:xt2_eql1},
  and
  \eqref{eq:error_asym2}.

\end{enumerate}

The above procedure is not always necessary.
As shown below, for a class of bandpass updating schemes,
  which resemble the Gaussian updating schemes,
  the asymptotically optimal schedule
  is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.

\subsection{\label{sec:cmpschemes}
Comparison of updating schemes}

Now that we can compute the optimal schedule
for a given updating scheme,
we will turn to the problem of finding
optimal updating schemes.

\subsubsection{\label{sec:optWL}
Asymptotic optimality of the single-bin scheme}

We will first show
that the single-bin scheme is asymptotically optimal.
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence the same $\Gamma_k$'s,
but with different positive eigenvalues,
$\lambda_k$'s.
Using the Cauchy-Schwarz inequality, we have,
for any set of nonnegative numbers, $c_k$'s\footnote{This
inequality follows from the nonnegativity of
the quadratic polynomial,
$$
\int_0^T
  dt \sum_{k = 1}^{n-1} \Gamma_k \,
    \left( {\dot \vartheta}_k \, \chi - c_k \right)^2
  \equiv
  A_2 \, \chi^2 + A_1 \, \chi + A_0
  ,
$$
which implies a nonpositive discriminant,
$A_1^2 - 4 \, A_0 \, A_2$.}
\begin{align}
&
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, {\dot \vartheta}_k^2\bigl( q(t) \bigr)
\right)
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k^2
\right)
\notag
\\
&
\qquad \qquad
\ge
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k \, {\dot \vartheta}_k \, \bigl( q(t) \bigr)
\right)^2
\notag
\\
&
\qquad \qquad
=
\left(
  \sum_{k = 1}^{n-1} \Gamma_k \, c_k
    \left[
      1 - e^{ -\lambda_k \, q(T) }
    \right]
\right)^2.
\notag
\end{align}
Note that the last expression %of Eq. \eqref{eq:CSineq}
depends on the curve, $q(t)$ ($0 < t < T$),
only through the endpoint value, $q(T)$,
which is fixed in the variational process.
Thus, the inequality sets a lower bound
for the asymptotic error $\Err_A(T)$
given in Eq. \eqref{eq:error_asym}.
The equality is achieved
if $\dot \vartheta_k\left( q(t) \right) = c_k$
for all $k > 0$ at any $t$,
up to a multiplicative factor.
Solving these equations yields
a set of identical eigenvalues,
$\lambda_1 = \cdots = \lambda_{n-1}$
and identical $c_k$'s.

Since an updating matrix
with identical eigenvalues
is a multiple of the identity matrix,\footnote{Since
  the eigenmode of $k = 0$ represents
  a uniform shift of the bias potential,
  we can freely set $\lambda_0$ to be
  the same as $\lambda_1$
  without changing the nature of
  the updating scheme.}
we recover the single-bin scheme
by setting the eigenvalues to $1$.
Thus, the single-bin scheme is most efficient
in error reduction in the asymptotic limit.
In particular,
with $\vartheta_k = (t+t_0)/(T+t_0)$,
we can readily recover the optimal schedule,
Eq. \eqref{eq:alpha_invt1}, with
the asymptotic error being
$\Err_A(T) = \sum_{ k = 1 }^{n-1} \Gamma_k \, T / (T + t_0)^2$,
which recovers Eq. \eqref{eq:EA_sbin1}.\footnote{By
Eq. \eqref{eq:parseval}, we have
$\sum_{k=1}^{n-1} \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
= \sum_{i=1}^n \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle$,
and by summing $t$ from $-\infty$ to $\infty$, we get
$\sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma$.}

\subsubsection{\label{sec:bandpass}
Bandpass updating schemes}

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that completely eliminates certain short-wavelength eigenmodes,
which are typically local noise.
A bandpass updating scheme is essentially
the single-bin updating scheme
accompanied by a perfect filter
for the designated noise modes,
which are assigned with zero eigenvalues.
For example,
by assuming
that the PMF is so smooth
that it can be represented by the first $K$
modes,\footnote{As an estimate of error,
  we note that truncating a linear or quadratic function
  at mode $K$ (for a non-periodic variable)
  results in an error roughly proportional to
  $|\max_i u_i - \min_i u_i|^2/K^3$.}
we demand
\begin{equation}
  \lambda_1 = \cdots = \lambda_{K-1} = 1,
  \mathrm{\; and \;}
  \lambda_K = \cdots = \lambda_{n-1} = 0.
  \label{eq:lambda_bandpass}
\end{equation}
The asymptotic error is then
$\Err_A(T) = \sum_{ k = 1 }^{ K - 1 } \Gamma_k T/(T + t_0)^2$,
which is less than that from Eq. \eqref{eq:EA_sbin1}.
The optimal schedule of these schemes
remains the inverse-time schedule
given by Eq. \eqref{eq:alpha_invt1}
with $t_0$ given by Eq. \eqref{eq:t0_sbin}.

Since bandpass updating schemes
straightforward generalizations of
the single-bin updating scheme,
we can also borrow the WL algorithm
to control the updating magnitude
in initial stages of the simulation.
However, the histogram fluctuation should
be limited to first $K$ modes,
as the bandpass updating scheme
only partially flattens these histogram
modes (cf. Appendix \ref{sec:hfluc}).

In addition to the application to homogeneous updating schemes
(Appendix \ref{sec:homo_bandpass}),
bandpass updating schemes can also be applied to
a continuous variable, $z$,\footnote{
  In the transition to a continuous variable,
  we will map
  $\rho_i$ to $\rho(z) \, \delta z$,
  $\phi_{ki}$ to $\phi_k(z) \, \delta z$,
  $w_{ij}$ to $w(z, z') \, \delta z$, etc.}
with the eigenmodes derived from orthogonal polynomials, $R_k(z)$,
that satisfy
\begin{align*}
  \int R_k(z) \, R_{k'}(z) \, \rho(z) \, d z = \delta_{k, k'}
  .
\end{align*}
With $\phi_k(z) = R_k(z) \, \rho(z)$,
we can readily satisfy Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0},
and construct the updating matrix by
Eqs. \eqref{eq:w_from_phi} and \eqref{eq:lambda_bandpass} as
\begin{align*}
  w(z, z')
  =
  \sum_{k=1}^{K-1} \frac{ \phi_k(z) \, \phi_k(z') } { \rho(z) }
  = \sum_{k=1}^{K-1} R_k(z) \, R_k(z') \, \rho(z'),
\end{align*}
where we have dropped the $k=0$ mode
that corresponds to a uniform shift
of the bias potential.
The resulting bias potential
assumes a polynomial form
and can be written as a superposition
$u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z)$,
with the coefficients updated as, according to Eq. \eqref{eq:mbin_update},
\begin{equation}
  c_k(t+1) = c_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr)
  .
  \label{eq:ckupdate}
\end{equation}

For example,
to sample a flat distribution, $\rho(z)$,
over $[z_c - \Delta z/2, z_c + \Delta z/2]$,
we can use the Legendre polynomials\cite{arfken},
and with $K = 2$,
the bias potential contains only a linear mode for
$R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
with its coefficient, $c_1$, updated according to
the difference between the average of $z$
and the center, $z_c$.
We thus recover the LLR
algorithm\cite{langfeld2012},
and the inverse-time schedule
was recommended for optimal convergence of $c_1$\cite{pellegrini2014}.

Alternatively,
to sample a Gaussian distribution
centered at $z_c$ with width $\sigma_z$
we can use the Hermite polynomials\cite{arfken}.
With $K = 3$, the bias potential is quadratic
and is a linear combination of
$R_1(z) = (z - z_c)/\sigma_z$
and
$R_2(z) = \bigl[(z - z_c)^2 - \sigma_z^2\bigr] /\left(\sqrt 2 \, \sigma_z^2\right)$.
Note that quadratic bias potentials are commonly
employed\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007}
in umbrella sampling\cite{torrie1974, *torrie1977}
and the method of Gaussian ensemble\cite{hetherington1987,
*challa1988, *costeniuc2006},
which is helpful in overcoming barriers
in first-order phase transitions.
By Eq. \eqref{eq:ckupdate},
$c_1$ and $c_2$ are updated
according to the first and second
central moments of $z$, respectively.
Below we will refer to this application
as adaptive Gaussian umbrella sampling.

The use of a polynomial basis makes
the bias potential smooth and differentiable,
and hence suitable for MD applications.
However,
the distribution, $\rho_i$,
has to be sufficiently narrow
because of the limited number of modes
in controlling the moments.
We can nonetheless sample a wider range
by constructing a mixture\cite{shirts2017}
or an expanded ensemble
from a superposition of multiple Gaussian umbrellas
centered at different places via
the frameworks of
parallel\cite{swendsen1986,
  *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
  *earl2005, *zuckerman2011, *rauscher2009, kim2010}
and/or simulated tempering\cite{marinari1992,
  *lyubartsev1992, li2007,
  park2007, *nguyen2013, *zhang2015st}.
While the mutual exchange probability
can be readily derived in the former
framework\cite{neuhaus2006, *neuhaus2007, kim2010},
the transition probability in the latter case
requires estimation of the partition function
of the Gaussian ensemble
$Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$
as the bias potential for inter-umbrella transitions.
Here,
$\beta$ is the inverse temperature,
$\mathcal U(\mathbf x)$ is the unbiased potential energy
as a function of molecular coordinates, $\mathbf x$,
and $z$ has also been written as a function of $\mathbf x$.
The transition probability from
the $i$th umbrella to the $j$th is given by
\begin{align*}
A(i\to j) =
\min\left\{1,
  \frac{Z^{(i)}}{Z^{(j)}}
  e^{
    u^{(i)}\bigl( z(t), t \bigr)
    - u^{(j)}\bigl( z(t), t \bigr)
    }
  \right\}
  ,
\end{align*}
where $z(t)$ is the current value of $z$.
While the asymptotic updating magnitude
would be given by the inverse-time formula,
we can employ the WL strategy early in simulation
to control the updating magnitude in stages
with the histogram fluctuation
redefined by Eq. \eqref{eq:maxFk}
to encompass both intra-umbrella and inter-umbrella
components.
After simulation,
the weighted histogram analysis method
(WHAM)\cite{ferrenberg1988, *ferrenberg1989,
*kumar1992, *roux1995, *bartels1997, *souaille2001,
*kobrak2003, *gallicchio2005, *chodera2007,
*fenwick2008, *kim2011, *shirts2008,
zhu2012, newman, frenkel},
can be used to recover the unbiased distribution.

\section{\label{sec:results}
Numerical results}

\subsection{\label{sec:lj}
Comparison of updating schemes}

Our first test system is
a Lennard-Jones (LJ) fluid of $100$ particles,
and we wish to compute the PMF along
the distance, $r$, between two special particles.
The two special particles are mutually non-interacting
but they interact normally with other particles,\footnote{The
two special particles were placed on an axis parallel to a side of the cubic box}
and the intrinsic distribution, $p(r)$, is
the cavity distribution function, $y(r)$\cite{hansen}.
We will assume reduced units in this calculation.
The reference bias potential was pre-calculated from a long simulation.

We used the Metropolis algorithm\cite{frenkel, metropolis1953}
for configuration sampling.
Each step consists of
one MC trial for displacing one of the two special particles,
and another four trials for displacing the other particles.
In each trial,
a random particle is displaced in each dimension
according to a uniform distribution over $(-\varepsilon^*, \varepsilon^*)$.
We used $\varepsilon^* = 0.1/\varrho^*$
to ensure a reasonable acceptance ratio,
and $\varrho^*$ is the reduced number density.

The PMF was calculated from $r = 0$ to half of the box size
(also the cutoff distance of the potential)
on a grid of size $\delta r = 0.01$ in reduced units.
Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
and the numbers of bins were $n = 500$ and $250$, respectively.
The reduced temperature was $T^* = 3$.
We tested the single-bin, Gaussian, and bandpass updating schemes.
Each simulation started with an ``equilibration'' phase
of $10^7$ steps under a constant updating magnitude,
$a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
was $0.05$ and $0.025$, respectively.}
followed by a production phase of another $10^7$ steps
with updating magnitude given by
either the inverse-time or optimal schedule.
We repeated each simulation multiple times
and report the average.
To make sure the same schedule for multiple runs under the same condition,
the optimal schedule was pre-computed from a set of fixed
$\Gamma_k$'s and $\epsilon_k$'s,
although ideally they should be estimated from the data
collected from the equilibration phase in individual runs.

As shown in Table \ref{tab:lj_error},
all updating schemes
with either the inverse-time or optimal schedule
were able to reduced
the error significantly,
and the errors roughly matched the predicted values
calculated using the pre-computed
$\Gamma_k$'s and $\epsilon_k$'s.
The Gaussian updating scheme produced less error than
the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case.
The effectiveness of the Gaussian updating scheme
in the low-density case
was probably because the sampling along
the collective variable was free from major barriers
and thus the histogram fluctuation largely
came from local noise.
In this case, we could further reduced the error
by using the bandpass updating scheme
(with the mode cutoff $K = 20$).
On the other hand, in the high-density case,
the error was dominated by the $k=1$ eigenmode
corresponding to the longest range fluctuation,
$\Gamma_1 \approx \Gamma$,
and the reduction of local noise
had little effect on the total error.

\begin{table}[h]\footnotesize
  \caption{\label{tab:lj_error}
    Error of the bias potential under
    the single-bin,
    Gaussian (labeled by the width, $\sigma$),
    and bandpass (labeled by the mode cutoff, $K$)
    updating schemes
    on the LJ system.
    The numbers have been averaged over multiple runs,
    and the predicted values are shown in parentheses for comparison;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase under a constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.1$,
      $500$ bins,
      $\Gamma_1 \approx 111$,
      $\Gamma \approx 1.1\times10^3$.
    } \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.2(11.4)\times10^{-5}$
    \\
    $\sigma=0.1$ & $1/t$
    & $7.5(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    $\sigma=0.1$ & opt.
    & $7.5(8.9)\times10^{-3}$
    & $3.3(3.9)\times10^{-5}$
    \\
    $\sigma=0.2$ & $1/t$
    & $6.6(7.8)\times10^{-3}$
    & $3.9(4.8)\times10^{-5}$
    \\
    $\sigma=0.2$ & opt.
    & $6.7(7.8)\times10^{-3}$
    & $2.4(2.9)\times10^{-5}$
    \\
    $\sigma=0.5$ & $1/t$
    & $5.4(6.6)\times10^{-3}$
    & $3.3(3.8)\times10^{-5}$
    \\
    $\sigma=0.5$ & opt.
    & $5.8(6.6)\times10^{-3}$
    & $2.7(3.0)\times10^{-5}$
    \\
    $K=20$ & $1/t$
    & $8.0(8.9)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.8$,
      $250$ bins,
      $\Gamma_1 \approx 2.0\times10^4$,
      $\Gamma \approx 2.7\times10^4$
    } \\
    \hline
    single-bin & $1/t$
    & $1.4(1.4)$
    & $3.4(2.7)\times10^{-3}$
    \\
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $3.0(2.8)\times10^{-3}$
    \\
    $\sigma=0.1$ & opt.
    & $1.2(1.3)$
    & $3.0(2.8)\times10^{-3}$
    \\
    $\sigma=0.2$ & $1/t$
    & $1.3(1.2)$
    & $3.5(3.2)\times10^{-3}$
    \\
    $\sigma=0.2$ & opt.
    & $1.2(1.2)$
    & $3.1(2.9)\times10^{-3}$
    \\
    $\sigma=0.5$ & $1/t$
    & $1.0(0.9)$
    & $9.9(8.0)\times10^{-3}$
    \\
    $\sigma=0.5$ & opt.
    & $1.0(0.9)$
    & $5.9(5.4)\times10^{-3}$
    \\
    $K=20$ & $1/t$
    & $1.4(1.3)$
    & $3.1(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

For the Gaussian updating scheme,
there was an optimal width for maximum effectiveness.
The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
or $0.07$ for $\varrho^* = 0.8$,
and it also depended on the simulation length
and other parameters.
Although widening the Gaussian can
effectively suppress random error
from local histogram fluctuation,
it also increased the risk of including
more systematic bias in middle-range modes,
which can be difficult to remove
even with the optimal schedule.
This is because the rate of error reduction
is proportional to the eigenvalue, $\lambda_k$
[cf. Eq. \eqref{eq:vt_diffeq_mbin}],
which decreases rapidly with the mode index $k$,
cf. Eq. \eqref{eq:lambda_Gaussian}.

To illustrate the mode dependence of the error,
we show in Fig. \ref{fig:lj_xerr}
the error components, $\langle \tilde v_k^2 \rangle$,
against the eigenmode index, $k$,
for the $\varrho^* = 0.1$ case
with width $\sigma = 0.2$.
For modes $k=10$ to $25$,
the final errors from the Gaussian updating scheme
were greater than those from the single-bin updating scheme
although the initial values were less.
For the Gaussian updating scheme,
the inverse-time schedule tended to
reduce the updating magnitude too quickly
leaving a hump on the error profile
for these middle-range eigenmodes.
The optimal schedule improved over the inverse-time schedule
and produced a flatter profile for the final error.
As a result, the optimal schedule usually produced
less error than the inverse-time schedule
as shown in Table \ref{tab:lj_error}.
Although the gain of using the optimal schedule
was modest in the reported cases,
it would increase with the simulation length.\footnote{
  For example, for the $\varrho^* = 0.1$ case
  with width $\sigma = 0.2$,
  the predicted error of a simulation of $10^8$ steps
  was $2.5\times10^{-5}$ for the inverse-time schedule
  or $3.7\times10^{-6}$ for the optimal schedule.}

One way to overcome the systematic bias
in the Gaussian updating scheme
is to use the correction formula
based on histogram, Eq. \eqref{eq:uhatav}.
As shown in Fig. \ref{fig:lj_xerr},
even for the simulation using the non-optimal $1/t$ schedule,
the correction was able to reproduce
the flatter profile seen in the single-bin case.
However, this correction also introduced
some local noise in the large $k$ modes,
and it would be best used as a post-simulation
analysis tool for simulations
with detectable residual bias.
For example, for the simulation in the $\varrho^* = 0.8$ case
with $\sigma = 0.5$ and the inverse-time schedule,
the correction was able to lower
the error from $9.8\times10^{-3}$
(Table \ref{tab:lj_error}) to $3.3\times10^{-3}$.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_xerr.pdf}
  }
  \caption{
    \label{fig:lj_xerr}
    Error components from FDS simulations
    using the Gaussian updating scheme
    of width $\sigma = 0.2$
    on the LJ system at density $\varrho^* = 0.1$.
    The optimal schedule (opt.) was more effective
    in reducing the errors of middle-range eigenmodes
    than the inverse-time schedule ($1/t$).
    The ``corrected'' values were obtained from
    Eq. \eqref{eq:uhatt}.
    The lines are a guide to the eyes.
  }
\end{center}
\end{figure}

In Fig. \ref{fig:lj_alpha},
we show the optimal schedules in a few cases.
Since we have chosen a long equilibration period
for the preliminary run under the constant updating magnitude, $a_0$,
the initial magnitude was around
the ideal value of $a_0/2 = 5\times10^{-5}$
given by Eq. \eqref{eq:half_alpha0}.
But if the Gaussian were too wide,
e.g. if $\sigma = 0.6$,
the eigenvalues $\lambda_k$ of some modes
became too small to damp out the initial bias,
resulting in a much larger initial updating magnitude, $\alpha(0)$
[cf. Eq. \eqref{eq:opt_alpha0}].
In this case, the predicted schedule
would be unusable because the resulting deviation in the bias potential
was no longer small,
and one may have to either decrease the width
or to elongate the preliminary simulation
under the constant magnitude.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
  }
  \caption{
    \label{fig:lj_alpha}
    Optimal schedules of Gaussian updating schemes
    for the LJ system.
    The starting updating magnitude was usually
    around $a_0/2 = 5\times10^{-5}$
    as given by Eq. \eqref{eq:half_alpha0}.
    But an overly wide Gaussian window ($\sigma = 0.6$)
    led to a much larger updating magnitude.
  }
\end{center}
\end{figure}

\subsection{\label{sec:potts}
Adaptive Gaussian umbrella sampling}

As another proof of principle,
we tested the adaptive Gaussian umbrella sampling method
introduced in Sec. \ref{sec:bandpass}
on the two-dimensional $L\times L$ ten-state
Potts model\cite{wu1982, newman, wang2001, wang2001pre}
in the energy space.
We spread the Gaussian umbrellas
of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
with an even spacing of $L$.
For simplicity,
only transitions between neighboring umbrellas
were performed.
Initially, we set $c_1 = 1.4 \sigma_E$, and $c_2 = 0$.
For configuration sampling in each umbrella,
we used the Wolff algorithm\cite{wolff1989, newman}
(which is more efficient than the Metropolis algorithm)
at temperature $c_1/\sigma_E$
followed by a Metropolis step with
acceptance probability
$\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
whose average was no less than 92\%.
The updating magnitude started at $0.01$,
and it underwent the WL procedure
with the fluctuation threshold being $0.2$
[using the definition of Eq. \eqref{eq:maxFk}]
and the reduction factor being $1/2$
until it fell below
the value given the inverse-time schedule\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
This updating magnitude was
used to update the values of both $\ln Z^{(i)}$'s
for inter-umbrella transitions
and $c_1^{(i)}$'s and $c_2^{(i)}$'s
for intra-umbrella sampling.

The results for the $L = 64$ case
after $10^5 L^2$ steps are shown
in Fig. \ref{fig:pt_hist}.
The Gaussian histograms were evenly spaced with similar width as intended,
as shown in Fig. \ref{fig:pt_hist}(a).
By WHAM, we found the critical temperature at $0.70162$
in agreement with a previous study\cite{wang2001pre}.
The parameter, $c_1/\sigma_E$, roughly represents
the local or statistical temperature,
and it demonstrated
the backbending behavior\cite{kim2006, *kim2007, kim2010}
typical in a first-order phase transition,
as shown in Fig. \ref{fig:pt_hist}(b).
The second term of the bias potential,
$c_2 \, R_2(E) = \sqrt 2 \, c_2 \, [(E - E_c)^2/(2 \, \sigma_E^2) - \frac 1 2]$,
provides a quadratic restraint on the distribution width.
The coefficient, $\sqrt 2 \, c_2$, represents the amount of ``pressure''
needed to restore the target Gaussian,
$\rho(E) \propto \exp[-(E - E_c)^2/(2 \, \sigma_E^2)]$,
and a value over $1.0$ suggests
that the intrinsic distribution
(or the density of states here)
is locally convex or unstable,\cite{kim2010}
which was so in the critical region,
as shown in Fig. \ref{fig:pt_hist}(c).

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/pt_hist.pdf}
  }
  \caption{
    \label{fig:pt_hist}
    Adaptive Gaussian umbrella sampling
    on the $64 \times 64$ ten-state Potts model.
    (a) Normalized histograms (solid lines)
    and the reweighted distribution
    at the critical temperature (dashed line).
    (b) The parameter, $c_1/\sigma_E$, roughly corresponding to
    the statistical temperature, $\beta(E)$,
    demonstrated the backbending behavior.
    (c) The parameter, $\sqrt 2 \, c_2$, exceeds $1.0$
    in the critical region, indicating local convexity
    of the density of states.
  }
\end{center}
\end{figure}

\section{\label{sec:conclusion}
Conclusions and Discussions}

In conclusion,
we have proposed a method of computing
the optimal schedule of the updating magnitude
for a general class of free energy calculation methods,
which we refer to as adaptive flat-distribution-sampling (FDS) methods
that include the Wang-Landau (WL) algorithm and metadynamics.
Adaptive FDS methods
incrementally update a bias potential
along a collective variable
to offset the potential of mean force (PMF)
and thereby sample a flat distribution.
The optimal schedule delivers the fastest convergence
of the bias potential,
and thus helps minimize the effort
of computing the PMF.

The optimal schedule depends on the updating window function,
which can be limited to a single-bin (as often in the WL case),
or extended to several neighboring bins
taking a Gaussian shape (as often in the metadynamics case).
We can generally characterize
an updating scheme with an updating matrix,
with each column being the normalized
window function for the bin of the column.
Our method computes the optimal schedule from
the eigenvalues of the updating matrix,
which represent the relative rates of error reduction,
and the integrals of the autocorrelation functions,
which represent the intrinsic histogram fluctuation,
of the eigenmodes.
For the single-bin scheme with a set of uniform eigenvalues,
we confirmed that the optimal schedule
is given by the inverse-time formula,
Eq. \eqref{eq:alpha_invt1}.
For a general multiple-bin scheme,
including the Gaussian one,
the optimal schedule, having to balance
the different rates of error reduction of the eigenmodes,
is usually not given by a simple closed form,
but is implicitly given by an equation of motion
of a free particle with a position-dependent mass,
Eq. \eqref{eq:q_opt}.
In this case,
the optimal schedule and error
can be sensitive to the simulation length
and other parameters.

In particular, the Gaussian updating scheme is often
equipped with a set of rapidly diminishing eigenvalues.
While this feature helps suppress
the accumulation of random error and
maintain the smoothness of the bias potential,
it also slows down the reduction
of the initial bias, if any,
in the short- and middle-range eigenmodes.
Thus, one needs to carefully choose the width of the Gaussian
and the optimal schedule of decreasing the updating magnitude,
which sometimes needs to be fixed at a maximal acceptable level
for a long time before any reduction can happen.
The possible efficiency loss in
error reduction of the long- and middle-range modes
can often be compensated
by the histogram-based correction formula, Eq. \eqref{eq:uhatav},
for post-simulation analysis,
which ideally would remove the remaining bias
and recover a bias potential with precision
similar to that from the single-bin case.

In comparing different updating schemes,
we found that
the single-bin updating scheme was optimal
in error reduction
in the long time limit
without a priori assumptions
on the smoothness of the PMF.
This scheme can be generalized to
a class of bandpass updating schemes,
which are also asymptotically optimal,
but completely filter the components
in the noise modes.
The optimal schedule of the bandpass schemes
is always given by an inverse-time formula.
If the bandpass updating scheme was constructed
from orthogonal polynomials,
the updating scheme became a generalization
of the LLR algorithm\cite{langfeld2012, pellegrini2014},
and it could be used for parameter adjustment
in umbrella sampling with
quadratic\cite{neuhaus2006, *neuhaus2007, zhu2012}
or similar\cite{kim2010}
bias potentials.

This study has several limitations.
First, we have ignored the \emph{systematic}
error\cite{zhou2005, morozov2007, zhou2008}
caused by the updates in adaptive FDS methods.
This simplification is justified in the asymptotic regime:
when the updating magnitude is small
and the sampling is a quasi-equilibrium one\cite{
  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
the random error,
which is roughly proportional to
the square root of the updating magnitude\cite{
  zhou2005, morozov2007, zhou2008, bussi2006},
can easily outweigh
the systematic error,
which is roughly proportional to
the magnitude itself\cite{morozov2007}.
Second, by using the white noise approximation,
we have represented the autocorrelation function
of each eigenmode by a single number, $\Gamma_k$.
Although this assumption appeared to work well
in our test systems modeling long simulations,
it may break down for relatively short simulations
of glassy systems.
Finally, the adaptive Gaussian umbrella sampling
approach requires refinement to improve
its stability for general application.
For example, we find that with slow sampling processes,
the updating scheme might occasionally overly update
the parameter, $c_2$, leading to unreasonable values,
and the parameters used here might not be optimal.

\section{Acknowledgments}

We thank Dr. Y. Mei, J. Wang,
O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
Dr. C. Myers, and Dr. O. Samoylova
for helpful discussions.
We gratefully acknowledge the Robert A. Welch Foundation (H-0037),
the National Science Foundation (CHE-1152876)
and the National Institutes of Health (GM-037657)
for partial support of this work.
JM thanks support from the National Institutes of Health
(R01-GM067801, R01-GM116280),
the Welch Foundation (Q-1512),
and National Science Foundation (Grant PHY-1427654).
This research used computing resources of
the National Science Foundation XSEDE grid.

\appendix

\section{\label{sec:equilerr}
Runtime average interpretation
of the inverse-time schedule
}

The optimality of the inverse-time schedule
for the single-bin updating scheme
can be understood from the observation that
the schedule allows the bias potential
to coincide with a runtime average
of independent corrections.

Assuming that we can linearize the histogram correction
in Eq. \eqref{eq:vcorr_equil} as
$\ln x \approx x - 1$ % (for $x \approx 1$),
and apply it to the instantaneous histogram, $h_i(t)$,
the independent estimate of bias potential in each step
would be given by
\begin{equation}
  \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { \rho_i } - 1
  = u_i(t) + f_i(t) - 1
  .
  \label{eq:uhatt}
\end{equation}
Below we will use capital letters to denote
time averages, e.g. for variable $x$, we define
$X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)$.
Time averaging Eq. \eqref{eq:uhatt} yields
\begin{align}
  \hat U_i(t)
  &\approx
  U_i (t)
  + F_i(t) - 1
  \approx
  U_i(t)
  + \ln \frac{ H_i(t) } { \rho_i }
  .
  \label{eq:uhatav}
\end{align}
This equation can also be derived\footnote{From Eq. \eqref{eq:uhatt},
  we can construct an independent estimator
  of the unbiased distribution as
  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$.
  Upon time averaging, we get
  $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
from the self-healing umbrella sampling
method\cite{marsili2006},
and it serves as
a generalization of Eq. \eqref{eq:vcorr_equil}
for FDS simulations under a variable bias potential
and we can use it to improve the estimated PMF
after simulation.
In particular, for the single-bin updating scheme,
Eq. \eqref{eq:wl_update},
we have, up to a constant shift,
\begin{align}
  \hat U_i(t)
  &\approx
  \frac{1}{t} \sum_{\tau=1}^t
  \left[
    u_i(\tau) +
    \frac{ u_i(\tau + 1) - u_i(\tau) } { \alpha(t) }
  \right]
  ,
  \notag
  \\
  &\approx
  \Delta u_i
  +
  \sum_{\tau=1}^t
  \left[
    1 + \frac{1}{\alpha(\tau - 1)} - \frac{1}{\alpha(\tau)}
  \right]
  \frac{ u_i(\tau) } {t}
  ,
  \label{eq:Uhat_sum}
\end{align}
where
$\Delta u_i \equiv u_i(t+1) /[t \, \alpha(t)]
- u_i(1)/[t \, \alpha(0)]$.
With the inverse-time schedule, Eq. \eqref{eq:alpha_invt},
each term in the sum vanishes,
and if we further define $\alpha(0) \to \infty$,
then
$$
u_i(t+1) = \hat U_i(t),
$$
i.e. the bias potential in this case coincides with
the runtime average of the correction, $\hat u_i(t)$,
at every step, $t$.
On the other hand,
with a constant $\alpha(\tau) = a_0$ for $\tau = 0, 1, \dots$,
Eq. \eqref{eq:Uhat_sum} is approximately a plain average\cite{zhou2005}:
$\hat U_i(t) \approx U_i(t)$,
if we can assume an equilibrium condition
with $u_i(t+1) \approx u_i(1)$.

We can also cast other updating relations,
such as Eqs. \eqref{eq:wl_update}
and \eqref{eq:vkupdate},
as runtime averages,
but with a dynamic weight.
The weighted average
\begin{align*}
  X_\omega(t)
  =
  \frac{
    \omega(0) \, x(0) + \cdots + \omega(t) \, x(t)
  }
  {
    \Omega(t)
  }
  ,
\end{align*}
where
$\Omega(t) \equiv \omega(0) + \cdots + \omega(t)$
is the accumulated weight,
is equivalent to the recurrence relation
\begin{align*}
X_\omega(t) = X_\omega(t-1)
  + \frac{\omega(t)}{\Omega(t)}
  \Delta \, x(t)
,
\end{align*}
where
$\Delta x(t) \equiv x(t) - X_\omega(t-1)$
is the proposed change from the previous average
from the data point at step $t$.
The relative weight, $\omega(t)/\Omega(t)$,
can be identified as the updating magnitude,
\begin{equation}
  \alpha^*(t) \equiv
  \frac{ \omega(t) } {\Omega(t)}
  =
  1 - \frac{ \Omega(t-1) } {\Omega(t)}
  .
  \label{eq:alpha_from_Omega}
\end{equation}
The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
follows from the plain average,
$\omega(t) = 1$ (for $t > 0$),
and updating magnitudes above and below this value
correspond to runtime averages that favor future and past data,
respectively.
In the case of Eq. \eqref{eq:vkupdate},
we can identify
the bias potential $\tilde v_k(t)$
as $X(t-1)$,
$\tilde f_k(t)$ as $\Delta x(t)$
by Eq. \eqref{eq:uhatt},
and
the mode-$k$ updating magnitude,
$\alpha(t) \, \lambda_k$,
as $\alpha^*(t)$.
This mapping shows that
updating magnitude should not exceed $1.0$.
One way to satisfy this condition
in converting the results
from the continuous-time setup is
to use Eq. \eqref{eq:alpha_from_Omega}
with
\begin{align*}
  \frac{ \Omega(t) } { \omega(0) }
  = \prod_{\tau=1}^t \frac{1}{1- \alpha^*(\tau)}
\approx \exp\left(\int_0^t \alpha^*(\tau) \, d\tau \right).
\end{align*}

\section{\label{sec:hfluc}
Histogram fluctuation}

Here we wish to show that for an FDS simulation
with a nearly-perfect bias potential,
the histogram fluctuation
is roughly equal to the final error
from a long FDS simulation under the single-bin updating scheme
with the inverse-time schedule.
With a sufficiently accurate and stable bias potential,
the fluctuation of
the instantaneous histogram is dominated by the random part,
\begin{equation}
  f_i(t) \approx \zeta_i(t)
  ,
  \label{eq:xi_zeta}
\end{equation}
and the same holds for the Fourier transform,
$\tilde f_k(t) \approx \tilde \zeta_k(t).$
If we define the time average,
\begin{align*}
\tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
=\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
,
\end{align*}
for the histogram fluctuation of mode $k$,
then the total fluctuation is
\begin{align}
  F^2(T)
  &\equiv
  \sum_{k=1}^{n-1} \tilde F_k^2(T)
  =
  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
  =
  \sum_{i=1}^n
  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
  .
  \notag
\end{align}
From Eq. \eqref{eq:xi_zeta}, we have
\begin{align*}
\bigl\langle \tilde F_k^2(T) \bigr\rangle
&=
\frac{1}{T}
\sum_{t=-\infty}^\infty
\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
\\
&\approx
\frac{1}{T}
\sum_{t=-\infty}^\infty
\bigl\langle
  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
\bigr\rangle
=
\frac{ \Gamma_k } { T }
,
\end{align*}
and the total histogram fluctuation is
$\bigl\langle F^2(T) \bigr\rangle = \Gamma/T$.
Comparing this to Eq. \eqref{eq:Emin_sbin},
we may rephrase the optimality of the inverse-time schedule
as the inequality,
\begin{equation}
  \Err(T)
  \ge
  \bigl\langle F^2(T+t_0) \bigr\rangle
  .
  \notag
\end{equation}

The relation between error and histogram fluctuation
supports the WL criterion for stage transitions.
For a bandpass updating scheme,
we can define the histogram fluctuation
from the largest mode
\begin{align}
  \max_{1 \le k < K} \left| \, \tilde F_k(T) \, \right|
  ,
  \label{eq:maxFk}
\end{align}
which allows straightforward extension
to a continuous variable, $z$, with
$\tilde F_k(T) = \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)$.

\section{\label{sec:Gamma_measure}
Integrals of the autocorrelation functions of the eigenmodes
}

Here we give a method of measuring
the integrals of the autocorrelation functions, $\Gamma_k$'s,
in an adaptive FDS simulation
under a constant updating magnitude.
This method has the advantage of not requiring
explicit computation of
the autocorrelation functions of the eigenmodes.

The idea is to use a modification
of Eq. \eqref{eq:xt2_eql}
that avoids the problem of not
knowing the intrinsic distribution, $p_i$,
hence the potential shift,
$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
Since the shift,
$u_i(t) - v_i(t)$,
hence its Fourier transform,
$\tilde u_k(t) - \tilde v_k(t)$,
is a constant of time,
${\tilde u}_{k}$ and ${\tilde v}_{k}$
should share the same variance.
But since the ensemble average of ${\tilde v}_{k}$
is zero, we have
$\operatorname{var} {\tilde u}_{k} =
 \operatorname{var} {\tilde v}_{k} =
 \left\langle {\tilde v}_{k}^2 \right\rangle$,
and by using Eq. \eqref{eq:xt2_eql},
we get
\begin{equation}
  \operatorname{var} {\tilde u}_k
  =
  \frac{1}{2} \,
  a_0 \, \Gamma_k \, \lambda_k,
\label{eq:varXt}
\end{equation}
where the variance can be computed from trajectory averages.
In practice, Eq. \eqref{eq:varXt}
would fail if some eigenvalue, $\lambda_k$, is close to zero.
So it is best used for the single-bin updating scheme.

Alternatively,
if the propagation along the collective-variable is Markovian,
then in the long time limit, the evolution can be approximately
described by a transition matrix, $\mathbf A$, with
\begin{align*}
\langle h_i(t) \rangle = \rho_i,
\quad
\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
\quad \mbox{(for $t \ge 0$)}
.
\end{align*}
Further if the transition matrix satisfies detailed balance
with a set of eigenvectors, $\psi_{li}$,
that are orthonormal with respect to $\pmb\rho$
[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
\eqref{eq:eig_orthonormal_rows}]
then we have the decomposition
$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
with $\Lambda_l$ being the eigenvalue of mode $l$.
The autocorrelation function can then be computed as
\begin{align*}
\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\sum_{l = 0}^{n-1} \Lambda_l^t \,
\left(
  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
\right)^2
.
\end{align*}
In particular, if the eigenvectors, $\pmb \phi_k$
and $\pmb \psi_k$, coincide,
we would have
$\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\Lambda_k^t,$
and
\begin{equation}
  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  = \frac{1+\Lambda_k}{1-\Lambda_k}
  .
  \notag
\end{equation}

The transition matrix method
tends to underestimate
the first few $\Gamma_k$'s.
We will nonetheless use it as a model to
study the behavior of $\Gamma_k$
in a local sampling process.
If we borrow Eq. \eqref{eq:lambda_Gaussian}
for $\Lambda_k$,
with $\sigma_z$ interpreted as the average move size along $z$,
then
\begin{align}
\Gamma_k = \coth\left[
  \left(
    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
  \right)^2
  \right].
  \notag
\end{align}
For small $k$,
we have $\Gamma_k \propto \Delta z^2/k^2$,\cite{bussi2006}
which means that doubling the sampling range, $\Delta z$,
would make the histogram flattening
at least four times as difficult,
i.e. FDS methods are most effective
for short ranges\cite{wang2001, wang2001pre}.
For a modestly large $k$,
$\Gamma_k \approx 1$,
and by Eq. \eqref{eq:xt2_eql},
we would expect a roughly flat tail
for the single-bin updating scheme
as $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
cf. Fig. \ref{fig:lj_xerr}.

\section{\label{sec:more_wband}
Homogeneous updating schemes
}

Here, we give some mathematical details
on the homogeneous updating schemes.
By the translational invariance,
these schemes
satisfy detailed balance,
Eq. \eqref{eq:w_detailedbalance},
for the flat distribution, $\rho_i = 1/n$,
i.e. the updating matrix, $\mathbf w$,
is symmetric.
The updating matrices
share the same set of eigenvectors
(which here are sines and cosines of different frequencies),
and thus are completely determined
by the eigenvalues.
Further, the multiplication of $\mathbf w$
can be reduced to a convolution with the window function.
Thus,
there is a one-to-one correspondence between
the eigenvalues of $\mathbf w$
and the window function.

\subsection{\label{sec:wband_eig}
Eigenvalues and window function}

For a periodic variable\cite{dama2014},
the updating matrix, $\mathbf w$,
assumes the following form
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
\notag
\end{equation}
where
$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
will be referred to as a window function.
To satisfy Eq. \eqref{eq:w_sumj},
we impose the normalization
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  ,
\label{eq:mu_normalization}
\end{equation}
with $\mu_l = 0$ for $|l| > b$.
If $b = n/2$, $\mu_{-b}$ is removed
from the sum in Eq. \eqref{eq:mu_normalization}
to avoid double counting.
For simplicity, we also impose the symmetry
\begin{equation}
  \mu_i = \mu_{-i}
  .
\label{eq:mu_symm}
\end{equation}

To find the eigenvectors,
we define for a periodic variable, $\phi_i$,
the out-of-boundary values by
\begin{equation}
  \phi_i = \phi_{i \pm n},
\label{eq:phi_pbc}
\end{equation}
such that $i \pm n$ lies in between $1$ and $n$.
Then, the multiplication of the matrix, $\mathbf w$,
is equivalent to a convolution with the window function:
\begin{equation}
  \sum_{ j = 1 }^n
    w_{ij} \, \phi_j
  =
  \sum_{ j = 1 - n }^{ 2 \, n }
    \mu_{i - j} \, \phi_j
  =
  \sum_{ l = -b }^{ b }
    \mu_l \, \phi_{ i - l}
  .
\label{eq:wmul_to_convol}
\end{equation}
As a result, the characteristic equation,
$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
can be solved by discrete Fourier transform.
We may construct a set of orthonormal eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
compatible with the periodicity
required by Eq. \eqref{eq:phi_pbc},
\begin{equation}
  \phi^{(k)}_i
  =
  \phi_{ki}
  =
  \frac{ \sqrt{ 2 - \delta_{k,0} } } { n }
  \,
  {\cos \atop \sin}
  \left(
    \frac{ k \, i \, 2 \, \pi }
         {      n             }
  \right)
  ,
  \notag
\end{equation}
where the function takes the cosine form for $k \le n/2$,
or the sine form otherwise.
Then one can readily verify that the eigenvalues are given by
  Eq. \eqref{eq:wband_eigenvalue},
  and there is a two-fold degeneracy,
  $\lambda_{n - k} = \lambda_k$.

For a non-periodic variable,
we use the reflective boundary condition\cite{bussi2006},
and change the form of the updating matrix to
\begin{equation}
  w_{ij}
  =
  \mu_{ i - j }
  +
  \mu_{ i + j - 1 }
  +
  \mu_{ i + j - 2 n - 1 }
  ,
  \notag
  \label{eq:w_band_refl}
\end{equation}
where the last two terms on the right hand side,
representing two reflective mirrors at
$i = 1/2$ and $i = n + 1/2$,
are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
to the equilibrium distribution\cite{bussi2006}.
Again, the window function should satisfy
Eqs. \eqref{eq:mu_normalization}
and \eqref{eq:mu_symm},
but
the cutoff $b$ only needs to be less than $n$.

Note that the two reflective mirrors effectively
define a periodic variable of period $2 \, n$.
Thus, the eigenvalues can be borrowed from
Eq. \eqref{eq:wband_eigenvalue},
with a doubled period,
i.e. $g$ is changed to $2$ in the non-periodic case.

We define the out-of-boundary values as
\begin{equation}
  \phi_i
  =
  \begin{dcases}
    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
  \end{dcases}
\label{eq:phi_refl}
\end{equation}
such that the matrix multiplication is still equivalent to
a convolution as Eq. \eqref{eq:wmul_to_convol}.
Since Eq. \eqref{eq:phi_refl}
defines a periodic variable of period $2 \, n$
with a reflective symmetry around $i = 1/2$,
the eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
should share the same periodicity,
and be even functions about the same axis:
\begin{equation}
  \phi^{(k)}_i
  =
  \phi_{k i}
  =
  \frac{ \sqrt{ 2 - \delta_{k, 0} } }
       {             n              }
  \cos \left[
       \frac{ k \, \left( i - \frac 1 2 \right) \, 2 \, \pi}
            {             2 \, n                           }
       \right]
  ,
\notag
\end{equation}
and the eigenvalues are given by
  Eq. \eqref{eq:wband_eigenvalue}
  with $g = 2$.

We can recover the window function from the eigenvalues
by inversely transforming Eq. \eqref{eq:wband_eigenvalue},
\begin{equation}
  \mu_i
  =
  \frac { 1 } { g \, n }
  \sum_{ k = 0 }^{ g \, n - 1 }
  \lambda_k
  \cos \left(
       \frac{ k \, i \, 2 \, \pi }
            {      g \, n        }
  \right)
  .
\label{eq:mu_from_lambda}
\end{equation}
In the non-periodic case,
we have defined
$\lambda_k \equiv \lambda_{2 \, n - k}$
for $k = n + 1, \dots, 2 \, n - 1$,
as well as
\begin{align}
  \lambda_n
  =
  (-1)^{ n - 1 }
  \lambda_0
  +
  2 \, \sum_{ k = 1 }^{ n - 1 }
      (-1)^{n - k - 1} \, \lambda_k
  ,
\label{eq:lambdan}
\end{align}
to satisfy the constraint, $\mu_n = 0$.

\subsection{Gaussian updating scheme}

We define a Gaussian updating scheme
as an updating scheme with a
Gaussian-shaped window function.
This updating scheme is commonly
used in metadynamics.
Below we show its stability
in the continuous limit.

For $n \gg 1$,
we define an equivalent continuous variable, $z$,
over $[0, \Delta z]$
with bin size
$\delta z = \Delta z/(g \, n)$.
The variable is related to the bin index, $i$, as
$z = i \, \delta z$,
and
$\mu_i = \mu(z) \, \delta z$.
By assuming the largest bin cutoff, $b \approx g \, n/2$,
we can approximate Eq. \eqref{eq:wband_eigenvalue}
as an integral:
\begin{equation}
  \lambda_k
  =
  \int_{-\Delta z/2}^{\Delta z/2}
    \mu(z) \, \cos\left( \frac{ 2 \, \pi \, k \, z } { \Delta z} \right)
    \, d z,
  \notag
\end{equation}
with the normalization
$1 = \int_{-\Delta z/2}^{\Delta z/2} \mu(z) \, dz$.

If the width of the Gaussian window, $\sigma_z$,
is very small compared to $\Delta z$,
then we can extend the limits of the integrals
to infinity, and
\begin{equation}
  \mu(z)
  \approx
  \frac{            1            }
       { \sqrt{ 2 \pi } \sigma_z }
  \exp\left(
        -\frac{   z^2   }
              { 2 \, \sigma_z^2 }
      \right)
  .
\notag
\end{equation}
Then, from Eq. \eqref{eq:wband_eigenvalue},
we get Eq. \eqref{eq:lambda_Gaussian},
which shows that all eigenvalues are positive.
In practice, however, some eigenvalues can turn negative
because of the truncation and discretization.
We can maintain the stability of the updating scheme
by redefining the Gaussian scheme
starting with Eq. \eqref{eq:lambda_Gaussian}
and inversely computing the window function
using Eq. \eqref{eq:mu_from_lambda}.

\subsection{\label{sec:homo_bandpass}
Bandpass updating schemes}

For the bandpass updating schemes,
the eigenvalues are either $0$ or $1$ and
it is convenient to set $\lambda_0 = 1$.
We can use Eq. \eqref{eq:mu_from_lambda}
to find the window function.
For a periodic variable,
we will take into account the two-fold degeneracy
and modify Eq. \eqref{eq:lambda_bandpass}
by setting $\lambda_{n-K-1}, \dots, \lambda_{n-1}$
also to $1$.
Then, $\mu_l = \mathrm{sinr}(K, l, n)$ and $w_{ij} = \mu_{i-j}$,
where
$\mathrm{sinr}(K, l, n)
  =
  \sin
  \frac{ (2 K - 1) \, l \, \pi }
       {              n        }
  / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.

For a non-periodic variable,
we have $\lambda_n = (-1)^{n-K}$
from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
The updating matrix is given by
$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.\footnote{$w_{ij}$
  is free from the oscillatory term, $(-1)^{n-K+1}$,
  that appeared in the window function, since
  its contribution to $\mu_{i-j}$ in Eq. \eqref{eq:w_band_refl}
  is canceled by its contribution to either $\mu_{i+j-1}$
  or $\mu_{i+j-2n-1}$,
  whichever exists, of the opposite sign.}

\bibliography{simul}
\end{document}